---
title: "Introduction to Programming in R"
tutorial:
  version: 1.4
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

<style>
p.comment {
background-color: #EDBB99;
padding: 10px;
border: 1px solid black;
font-style: italic;
}
</style>

<style>
p.comment1 {
background-color: #A2D9CE;
padding: 10px;
border: 1px solid black;
}
</style>

<style>
p.comment2 {
background-color: #EDBB99; 
padding: 10px;
border: 1px solid black;
}
</style>

<style>
p.comment3 {
background-color: #FDFEFE; 
padding: 10px;
border: 1px solid black;
}
</style>


```{r setup, include=FALSE, context="setup"}
# devtools::install_github("rstudio/packrat")
library(tidyverse)
library(learnr)
library(gradethis)
library(checkr)
library(sortable)
library(here)

knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = grade_learnr)
tutorial_options(exercise.startover = TRUE)
tutorial_options(exercise.completion = FALSE)

new_recorder <- function(tutorial_id, tutorial_version, user_id, event, data) {
    #cat(tutorial_id, " (", tutorial_version, "): ", user_id, ", ", event, ", ", "\n", sep = "")
  cat(tutorial_id, " (", tutorial_version, "): ", user_id , "\n", sep = "")
  cat("event: ", event, "\n", sep = "")
}

options(tutorial.event_recorder = new_recorder)

is_book = FALSE
is_tutorial = TRUE
```

## Welcome

This tutorial will introduce you programming in the R programming language.

## Your first program

```{r, child=paste0(here(),"/src/snippets/r_programming_hello_world.Rmd")}
```

Once you have successfully edited the program, click on the "Next Topic" button below.


## Data

```{r, child=paste0(here(),"/src/snippets/r_programming_data.Rmd")}
```


## Operators

```{r, child=paste0(here(),"/src/snippets/r_programming_operators.Rmd")}
```


## Storing and reusing results with *variables*

```{r, child=paste0(here(),"/src/snippets/r_programming_variables.Rmd")}
```


## How R works

```{r, child=paste0(here(),"/src/snippets/r_programming_how_it_works.Rmd")}
```


## Boolean data

```{r, child=paste0(here(),"/src/snippets/r_programming_boolean_data.Rmd")}
```


## Vectors

```{r, child=paste0(here(),"/src/snippets/r_programming_vectors.Rmd")}
```


## Functions

```{r, child=paste0(here(),"/src/snippets/r_programming_functions.Rmd")}
```


## Packages

```{r, child=paste0(here(),"/src/snippets/r_programming_packages.Rmd")}
```


### The End!

That's all, folks!



