

Sometimes dataset creators start with grand intentions, but motivation or reality gets in the way. Because of this, the data at the top of datasets often looks great. However, the last rows in the dataset may contain incomplete or problematic data that got added on at the end.

We can examine both the first and last rows of the dataset with the `head` and `tail` functions respectively. The syntax of these functions is `head(dataset_name)`, e.g.:

```{r, echo = TRUE}
head(diamonds)
```

to display 6 rows from the top (or bottom) of the dataset. We can change the number of rows to display with the second argument:

```{r, echo = TRUE}
head(diamonds, 10)
```

We could also rewrite the previous code chunk using the pipe operator:

```{r, eval=FALSE, echo=TRUE}
diamonds %>%
  head(10)
```

<!-- Book version -->

<!-- TODO: 
book exercises on head() and tail()
-->

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Use the `tail` function to display the **last 8 rows** of the diamonds dataframe. Note that this code uses the pipe operator `%>%` to supply the first.
</p>

```{r cqimxllkpywvjfns, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Replace the blanks with either a function name, or a number
diamonds %>%
  _____(___)
```

```{r cqimxllkpywvjfns-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  tail(8)
```

```{r cqimxllkpywvjfns-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

As you can see, the `diamonds` data is actually pretty nice at both the top and the bottom (as you might hope for an "offical" ggplot dataset). The example datasets that come with packages are often nice like this - unfortunately real-world data is usually less friendly.