For a continuous variable, we can calculate a few statistics (a number representative of other numbers) to summarize the:

1. center (i.e. mean, or median), &

2. spread (e.g. the standard deviation, the range, the interquartile range, the minimum and maximum, etc.)

of a distribution.

An easy way to do this is with the `summarize()` function that we learned about in @sec-wrangling-groupby-summarize.

```r
dataframe %>%
  summarize(
    mean    = mean(continuous_variable),
    median  = median(continuous_variable),
    std_dev = sd(continuous_variable),
    iqr     = IQR(continuous_variable),
    min     = min(continuous_variable),
    max     = max(continuous_variable)
  )
```

<!-- TODO: 
comments on interpreting these statistics:
- center: are the mean and median similar to each other (especially relative to the spread), or do they indicate a skewed distribution? >median indicates a left skew, <median indicates a right skew.

- what is the std. dev., and how is it calculated.
- use vs the IQR.
- Things to look for with the min & max: outliers, weird values indicative of coding errors.

Also: use of `na.rm = TRUE`
-->

<!-- Book version -->

<!-- TODO: 
book exercises on continuous summary statistics
-->

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Calculate summary statistics the 6 summary statistics listed above for the continuous variable `carat` from the `diamonds` dataframe.
<p>')
```

```{r rsrzqnjqelsdcbxe, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Fill in the blanks:
diamonds %>%
  summarize(
    mean    = mean(____),
    median  = _______,
    std.dev = _______,
    ________,
    ________,
    ________
  )
```

```{r rsrzqnjqelsdcbxe-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  summarize(
    mean    = mean(carat),
    median  = median(carat),
    std.dev = sd(carat),
    iqr     = IQR(carat),
    min     = min(carat),
    max     = max(carat)
  )
```

```{r rsrzqnjqelsdcbxe-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

In our original question, we were interested in two variables: size (i.e. `carat`) and `price`. We should therefore calculate summary statistics for the `price` variable as well

<!-- Book version -->

<!-- TODO: 
book exercises continuous summary statistics of `price`
-->

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Repeat your summary statistics calculation for the continuous variable `price`.
<p>')
```

```{r cybpzfzntosxuqtr, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# One big blank to fill in this time!
diamonds %>%
  _______
```

```{r cybpzfzntosxuqtr-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"You should be able to just change the variable in you previous code from carat to price"
```

```{r cybpzfzntosxuqtr-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  summarize(
    mean    = mean(price),
    median  = median(price),
    std.dev = sd(price),
    iqr     = IQR(price),
    min     = min(price),
    max     = max(price)
  )
```

```{r cybpzfzntosxuqtr-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->


<!-- TODO: 
convert to single summarize here for univariate summary statistics. Move group_by() to covariation section 
-->

If a variable is categorical, we can summarize it by looking at the proportions of observations in different categories. To do this we will first use the `group_by()` to create groups of the same variable that we want to summarize:

```r
dataframe %>%
  group_by(categorical_variable) %>%
  summarize(
    count = n(),
    proportion = n() / nrow(.),
    percentage = 100 * proportion
  )
```

Notes on this code:

  * the `n()` function returns the number of rows *within each group*. It does not need an argument.
  
  * the `nrow()` function also calculates the number of rows, but needs an argument (a variable whose rows it should count).
    
    * the `.` argument inside `nrow` is used to refer to the original `dataframe` that we piped in.
    
    Thus `nrow(.)` will give us the total number of rows in the original dataframe (ignoring the groups that we created).
    
<!-- TODO: 
Interpretation of categorical summary statistics
- 
-->

<!-- Book version -->

<!-- TODO: 
book exercises categorical summary statistics
-->

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Calculate summary statistics for the categorical `cut` column of the `diamonds` dataframe.
</p>')
```

```{r lwsyhxxvndkapsgg, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Replace the blanks, following the template above
diamonds %>%
  group_by(____) %>%
  _____(
    count = n(),
    proportion = ______,
    _____ = _______
  )
```

```{r lwsyhxxvndkapsgg-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"You will need to fill in the categorical variable to group by, the `summarize` function, and the three expressions inside the summarize function"
```

```{r lwsyhxxvndkapsgg-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  group_by(cut) %>%
  summarize(
    count = n(),
    proportion = n() / nrow(.),
    percentage = 100 * proportion
  )
```

```{r lwsyhxxvndkapsgg-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

