<!-- TODO: start with single summarize here for univariate. Move group_by() to covariation section -->

If a variable is categorical, we can summarize it by looking at the proportions of observations in different categories:

```r
dataframe %>%
  group_by(categorical_variable) %>%
  summarize(
    count = n(),
    proportion = n() / nrow(.),
    percentage = 100 * proportion
  )
```

Notes on this code:

  * the `n()` function returns the number of rows. It does not need an argument.
  * the `nrow()` function also calculates the number of rows, but needs an argument (a variable who's rows it should count).
  * the `.` argument inside `nrow` is used to refer to the original `dataframe` that we piped in.

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Calculate summary statistics for the categorical `cut` column of the `diamonds` dataframe.
</p>

```{r lwsyhxxvndkapsgg, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Replace the blanks, following the template above
diamonds %>%
  group_by(____) %>%
  _____(
    count = n(),
    proportion = ______,
    _____ = _______
  )
```

```{r lwsyhxxvndkapsgg-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"You will need to fill in the categorical variable to group by, the `summarize` function, and the three expressions inside the summarize function"
```

```{r lwsyhxxvndkapsgg-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  group_by(cut) %>%
  summarize(
    count = n(),
    proportion = n() / nrow(.),
    percentage = 100 * proportion
  )
```

```{r lwsyhxxvndkapsgg-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

For a continuous variable, we typically use a few statistics (a number representative of other numbers) to summarize the:

1. center (i.e. mean, median, mode), &

2. spread (e.g. the standard deviation, the range, the interquartile range, etc.)

of a distribution.

```r
dataframe %>%
  summarize(
    mean    = mean(continuous_variable),
    median  = median(continuous_variable),
    std.dev = sd(continuous_variable),
    iqr     = IQR(continuous_variable),
    min     = min(continuous_variable),
    max     = max(continuous_variable)
  )
```

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Calculate summary statistics the 6 summary statistics listed above for the continuous variable `carat` from the `diamonds` dataframe.
<p>

```{r rsrzqnjqelsdcbxe, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Fill in the blanks:
diamonds %>%
  summarize(
    mean    = mean(____),
    median  = _______,
    std.dev = _______,
    ________,
    ________,
    ________
  )
```

```{r rsrzqnjqelsdcbxe-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  summarize(
    mean    = mean(carat),
    median  = median(carat),
    std.dev = sd(carat),
    iqr     = IQR(carat),
    min     = min(carat),
    max     = max(carat)
  )
```

```{r rsrzqnjqelsdcbxe-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

In our original question, we were interested in two variables: size (i.e. `carat`) and `price`. We should therefore calculate summary statistics for the `price` variable as well

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Repeat your summary statistics calculation for the continuous variable `price`.
<p>

```{r cybpzfzntosxuqtr, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# One big blank to fill in this time!
diamonds %>%
  _______
```

```{r cybpzfzntosxuqtr-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"You should be able to just change the variable in you previous code from carat to price"
```

```{r cybpzfzntosxuqtr-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  summarize(
    mean    = mean(price),
    median  = median(price),
    std.dev = sd(price),
    iqr     = IQR(price),
    min     = min(price),
    max     = max(price)
  )
```

```{r cybpzfzntosxuqtr-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->