Visualization is a vital part of exploratory data analysis and we need to pair it with summary statistics to make sure that we are not being fooled by those statistics.

Consider Anscombe's quartet, a series of 4 small datasets (each has 11 observations of an *x* and a *y* variable).

Here's one of the datasets:

```{r}
anscombe %>%
  select(x1, y1) %>%
  head(11)
```

Here's another:

```{r}
anscombe %>%
  select(x2, y2) %>%
  head(11)
```

The numbers seem pretty similar, right? In fact, the summary statistics are identical:

```{r}
anscombe %>%
  select(x1:y4) %>%
  mutate(index = row_number(x1)) %>%
  gather(x1:y4, key="column", value="value") %>%
  group_by(column) %>%
  summarize(
    mean = mean(value),
    standard.deviation = sd(value)
  )
```

The mean (center) and standard deviation (variation) is the same for all 4 *x*'s, and all 4 *y*'s.

But, if we plot the four distributions...

```{r}
anscombe %>%
  mutate(index = row_number(x1)) %>%
  gather(x1:y4, key="column", value="value") %>%
  separate(column, into = c("dimension", "dataset"), sep=c(1)) %>%
  pivot_wider(names_from=dimension, values_from=value) %>%
  ggplot() +
  geom_point(aes(x,y)) +
  facet_wrap(~dataset, ncol=2)
```

Visualization allows us to detect:

* outliers (points that lie away from the rest)
* characterize relationships between variables: are they linear (straight) or non-linear (curved)?
* identify hypotheses (theories) about the data
* spot problems that might exist in the data

<!-- Book version -->

<!-- TODO:
Anscombe's quartet book exercises
-->

<!-- End book -->

<!-- Tutorial version -->

```{r anscombe_outliers_quiz, eval=is_tutorial, echo=FALSE}
question("Which of Anscombe's quartets has outliers?",
  answer("1"),
  answer("2"),
  answer("3", correct = TRUE),
  answer("4", correct = TRUE),
  allow_retry = TRUE
)
```

```{r anscombe_linear_quiz, eval=is_tutorial, echo=FALSE}
question("Which of Anscombe's quartets show a non-linear relationship between $x$ and $y$?",
  answer("1"),
  answer("2", correct = TRUE),
  answer("3"),
  answer("4"),
  incorrect = "When assessing linearity, look at the overall pattern of the majority of data points. One or two points in strange places could just be outliers from an otherwise linear pattern.",
  allow_retry = TRUE
)
```

```{r anscombe_no_relationship_quiz, eval=is_tutorial, echo=FALSE}
question("Do any of Anscombe's quartets have data that might indicate no relationship between $x$ and $y$?",
  answer("1"),
  answer("2"),
  answer("3"),
  answer("4", correct = TRUE, message = "4 could have no relationship if the outlier is a error (rather than a correct observation."),
  incorrect = "One of the quartets has no relationship between $x$ and $y$ for most of the data points, except for one outlier...",
  allow_retry = TRUE
)
```

<!-- End tutorial -->
