With a mix of continuous and categorical variables, the general strategy is to graph the continuous variable(s) and then break the graph down into separate parts by the categorical variable(s).

For example, say you had one continuous variable and one categorical variable. We can graph the single continuous variable, using one of the graph types appropriate for a single continuous variable such as a histogram, box plot, or violin plot, and then break that graph into different pieces using the categorical variable.

For example, let's say we want to visualize the relationship between `color` and `price` of our diamonds. One way to break our graphs down in discrete pieces is using the `fill` aesthetic to add colors to the inside of our shapes:

```{r, echo=FALSE, fig.show='hide'}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = color), bins = 15)

ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = price, fill = color))

ggplot(data = diamonds) +
  geom_violin(mapping = aes(x = price, y="", fill = color))
```

```{r, echo=FALSE}
h <- ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = color), bins = 15)

b <- ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = price, fill = color))

v <- ggplot(data = diamonds) +
  geom_violin(mapping = aes(x = price, y="", fill = color))

ggarrange(h,b,v, ncol = 3, nrow = 1, common.legend = TRUE)
```

Note that colors are randomly assigned by `ggplot` (i.e. these are not the actual colors of the diamonds). We can see, however, that there does seem to be a small effect of color on price: although most of the distributions overlap, the center of some color categories occurs at higher prices than others. This leads to a strange conclusion: the diamond color categories range from the best coloring ("D") to the worst coloring ("J"), just from our graphs above it almost looks like the worst colored diamonds have higher prices.

The graphs above using the previously ignored y-axis of the box plot and violin plot to create separate boxes and violins. We can actually achieve the same effect without color^[no-color-bw] by mapping the categorical variable to the `y` aesthetic instead of the `fill` aesthetic:

^[no-color-bw]: Useful if you are restricted to black and white.

```{r, echo=FALSE}
b <- ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = price, y = color))

v <- ggplot(data = diamonds) +
  geom_violin(mapping = aes(x = price, y = color))

ggarrange(b,v, ncol = 2, nrow = 1)
```

Unfortunately we cannot do this with our histogram because it already uses the y-axis for the count. The other problem with the default colored histogram is that each color is stacked on top of the colors beneath it. Unfortunately this violates one of the principles of visualization that we previously learned about, which is that if we are using length (i.e. height of bins) to communicate information, then we ideally want those lengths to start at the same point. If, as in the histogram above, the bins all start in different positions, then we have to mentally try to move the bins, and then it becomes very hard to figure out what is going on. For example it is almost impossible to see the distribution differences in the histogram that are readily apparent in the box plots and violin plots.

We have two strategies to fix this:

1. We can add the `position = "identity"` argument to the `geom_histogram()` function to start all the colored bins from the x-axis, and then because the different colors will all overlap we will also need to add the `alpha` parameter to make them transparent.

2. Instead of solid bars, we can just connect the top of each bin with lines. This type of graph is called a *frequency polygon*, and can be created with the `geom_freqpoly()` geom function.

```{r}
h <- ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = color), bins = 15)

ho <- ggplot(data = diamonds) +
  geom_histogram(
    mapping = aes(x = price, fill = color), 
    bins = 15,
    position = "identity",
    alpha = 0.3
    )

fp <- ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = price, color = color), bins = 15)

ggarrange(h,ho,fp, ncol = 3, nrow = 1, common.legend = TRUE)
```

As we can see, the overlapping histogram has so many colors that it is a muddy mess! In my experience this muddle of colors is unavoidable if you have more than two categories, so colored overlapping histograms should usually only be used if you have two categories (or if you have categories that do not overlap much along the x-axis). By comparison, the difference between diamond types are definitely easier to spot in the frequency polygon.

This frequency polygon also makes it clear why the *J* colored diamonds seemed to be more expensive in the violin and box plots: we just don't have many small (and cheaper) poorly colored diamonds. Either such diamonds don't exist or, more likely, they were not included in this dataset because nobody cares about diamonds that are both small and ugly.

#### Faceting

Another strategy for breaking our graph up by a categorical variable is to split it into sub-plots. In `ggplot` speak this is called "faceting". We do this by adding an additional faceting function to the end of our graph code.

We can add the `facet_wrap()` function to any ggplot graph to facet over a single categorical variable. This will create a separate sub-plot ("facet") for each category in that categorical variable. For example, to facet our histogram of diamond prices over the `cut` variable:

```{r}
diamonds |>
  ggplot() +
  geom_histogram(mapping = aes(x = price), bins = 15) +
  facet_wrap(~ cut)
```

Note how we passed the argument `~ cut` to `facet_wrap()` to tell it what categorical variable to facet over? You must include the squiggly `~` symbol before the faceting variable's name in the `facet_wrap()` function, otherwise you will get an error.

#### Adding more variables

To add more variables, you can combine the different aesthetics and functions. Typically each variable you want to visualize will need to be mapped to its own aesthetic or facet.

For example, if we wanted to visualize two continuous variables and two categorical variables, then we could first map the continuous variables to the x- and y-axes of a scatter plot, and then map one of the categorical variables to the color parameter and facet over the second categorical variable.

Just bear in mind that the more information you cram into a graph, the harder it will become to interpret. Here is a plot of the covariation between the `price`, `carat`, `color`, and `cut` variables from the `diamonds` dataframe:

```{r}
diamonds |>
  ggplot() +
  geom_point(mapping = aes(x = carat, y = price, color = color)) +
  facet_wrap(~ cut) +
  labs(
    title = "Effect of color, cut, and size on diamond price",
    x = "size (carats)",
    y = "price (USD)"
  )
```


