How can we boil the covariation of two or more variables down to a single number?[^covariation_stats]

^[covariation_stats]: We will think about this more in @sec-modeling-chapter.

With continuous variables, we can calculate a statistic that measure the correlation. There are various such *correlation coefficients*, but the key idea is that they indicate how much of the variation in different variables is shared.

One such number is Pearson's correlation coefficient, which varies between -1 and 1. 

<!-- TODO: add mathematical details to appendix -->

```{r}
set.seed(41)
df <- tibble(
  x = 1:10,
  `Perfect positive correlation` = x,
  `Perfect negative correlation` = 10-x,
  `Weak correlation` = runif(10,0,10)
)
cor_df <- correlate(df) |>
  select(name = term, coef = x) |>
  filter(name != "x")
df |>
  pivot_longer(cols = `Perfect positive correlation`:`Weak correlation`) |>
  mutate(
    name = factor(name, levels = c("Perfect positive correlation", "Weak correlation", "Perfect negative correlation"))
  ) |>
  ggplot() +
  geom_point(aes(x,value)) +
  geom_text(aes(x = 5, y = 0, label = paste("correlation =",round(coef,3))), data = cor_df) +
  facet_wrap(~name)
```

To calculate *pairwise* correlations between columns in a dataframe, we can use the `correlate()` function from the `corrr` package. For example, we can look at the correlations between the `carat`, `depth`, and `price` columns of the `diamonds` dataframe:

```{r}
diamonds |>
  select(carat, depth, price) |>
  correlate()
```

We can see that there is a strong positive correlation of about 0.92 between `carat` and `price` (unsurprisingly!), but a weak relationship between the other two pairs of variables. 

Note that each correlation is reported twice in the table above. We can remove these unnecessary duplications by piping this output dataframe of `correlate()` onto the `shave()` function (also from the `corrr` package), which will just keep one half of the correlations.

We can also make the output a little prettier by piping the output dataframe of `shave()` onto a third function called `fashion()`, which will round the long correlation coefficients for us to just two decimal places:

```{r}
diamonds |>
  select(carat, depth, price) |>
  correlate() |>
  shave() |>
  fashion()
```

<!-- 
ToDo: Exercises
-->

<!-- TODO: 
 - continuous & categorical: group_by then summarize (note that we cannot group_by a continuous variable...)
- categorical variables: counts (cross-tabulation?) and categorical correlation coefficients -> add mathematical details to appendix 
- possibly create a flow chart of options - i.e. what to do if you have X continuous variables and Y categorical?
-->
