

If you have two continuous variables, such as `price` and `carat` in the `diamonds` dataset, then a scatter plot is generally a good method of examining their covariation:

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Create a scatter plot with `price` on the y-axis and `carat` on the x-axis.
</p>

```{r kfxhpnqgmtclqbkt, exercise = TRUE, eval=is_tutorial, echo=FALSE}
diamonds %>%
  ggplot() +
  ______ +
  labs(title = "Scatter plot of dimond price vs size", y = "price ($)", x = "size (carats)")
```

```{r kfxhpnqgmtclqbkt-solution, eval=is_tutorial, echo=FALSE}
diamonds %>%
  ggplot() +
  geom_point(mapping = aes(x = carat, y = price)) +
  labs(title = "Scatter plot of dimond price vs size", y = "price ($)", x = "size (carats)")
```

```{r kfxhpnqgmtclqbkt-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

We can see several things about this graph:

* There's obviously a positive relationship between these two variables.
* It's hard to tell the density of the points (i.e. where most of the points fall), because in many places all the point overlap in a blog.

We have a several options to show density.

* We could make the points transparent with the `alpha` parameter. Then dense regions would be darker (because there would be many overlapping transparent points). For example, with `alpha = 0.1`:

  ```{r, echo = TRUE}
  diamonds %>%
    ggplot() +
    geom_point(mapping = aes(x = carat, y = price), alpha=0.1) +
    labs(title = "Scatter plot of dimond price vs size", y = "price ($)", x = "size (carats)")
  ```

* We can also create a heat map, which is a colored grid where the intensity of the color represents a value on a scale - in this case, the color represents the density. This uses the `geom_bin2d` geom function:

  ```{r, echo = TRUE}
  diamonds %>%
    ggplot() +
    geom_bin2d(mapping = aes(x = carat, y = price)) +
    labs(title = "Heat map of dimond price vs size", y = "price ($)", x = "size (carats)")
  ```
  
  This is essentially a 2-dimensional histogram. Each square is like a bar in a univariate histogram. In this heat map, the color represents the "height" of the 2-dimensional bin, i.e. how many diamonds fall into each cell of the heat map's grid.

<!-- Todo
Have students recreate this graph
-->
