

There are all kinds of other graphs we can plot.

In the code cells below, experiment on your own with some exploratory data analysis of the `diamonds` dataset.

From our previous explorations, it there is a positive relationship between `price` and `carat`; in other words, as one variable increases, so does the other. However:

* the relationship was not perfect: there seemed to be some uneven scatter. What other variables might be influencing this relationship?

* relatedly, what other variables are related to price (& what is the variation within those other variables by themselves)?

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

In the following code boxes, play around with some EDA of your own. You might find it helpful to look at the ["What graph should I plot" cheatsheet](https://drive.google.com/file/d/1zsedQ9kHFtxhFE3R99PQ1a_yTTamJu3e/view?usp=sharing) for inspiration, and remember that you can always Google a geom function (or ask an instructor) to get more information on how to use it.

```{r dmscqgizgqobassd, exercise = TRUE, eval=is_tutorial, echo=FALSE}


```

```{r iemrbivllnatvsyb, exercise = TRUE, eval=is_tutorial, echo=FALSE}


```

```{r jqfjswlmbfodoaaw, exercise = TRUE, eval=is_tutorial, echo=FALSE}


```

```{r efylzsvfvlntqbyz, exercise = TRUE, eval=is_tutorial, echo=FALSE}


```

<!-- End tutorial -->


