
<!-- TODO: 
ggplot labs() section was moved from EDA and needs updating to be consistent with chapter
-->

It is good practice to label all the graphs we create. We can do this by adding the `labs()` function to a graph:


```{r, echo=TRUE}
starwars %>%
  ggplot() +
  geom_point(
    mapping = aes(x = height, y = mass)
  ) +
  labs(
    title = "Mass vs. height of Star Wars characters", 
    y = "mass (kg)", 
    x = "height (cm)"
    )
```

Notice how:

* the `labs()` function is a separate function that we add on to the graph as a separate layer with the `+` operator
* inside the `labs()` function, we can supply an argument to the `title` parameter to change the title, and the `y` and `x` arguments to change the labels on those axes.
  * A good title should succinctly describe what is being plotted on the graph.
  * Axis labels should indicate what the variable is, and what units it is measured in.
  
We can also update any legend label as well by providing the same parameter name that we used in the `aes()` function. For example, with `fill`:

```{r}
ggplot(data = starwars) +
  geom_histogram(
    mapping = aes(x = height, fill = gender),
    position = "identity",
    alpha = 0.4
    ) +
  labs(
    title = "Histogram of heights of Star Wars characters",
    x = "height (cm)",
    fill = "Gender of character"
  )
```

* Note that I have left the default `count` label on the y-axis of this histogram, since that is a perfectly adequate name and has no units.
* In the title I have explicitly called out that this is a histogram, since there are other types of graphs that also use bars and so we can make it clear to any viewer what is going on here. (There was no need to do anything similar in the previous scatter plot, since it was clearly a scatter plot an is unlikely to be confused with a different type of graph.)

<!-- Book version -->

<!-- TODO: 
book exercises on labs()
-->

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Using the `labs()` function, add a title and y-axis label to the boxplot of `price` that you created earlier (we will leave the x-axis label as its default, since that is adequate for this graph.)
</p>

```{r boxplot_labels_exercise, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Fill in the blanks:
diamonds %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = "price", y = price)) +
  _____(_____ = "A boxplot of price", _____ = "price ($)")
```

```{r boxplot_labels_exercise-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = "price", y = price)) +
  labs(title = "A boxplot of price", y = "price ($)")
```

```{r boxplot_labels_exercise-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->
