
*TODO: This was moved from EDA and needs updating to be consistent with chapter*

It is good practice to label all the graphs we create. We can do this by adding the `labs()` function to a graph:

```{r, echo=TRUE}
diamonds %>%
  ggplot() +
  geom_violin(mapping = aes(x = "price", y = price)) +
  labs(title = "Violin plot of price", y = "price ($)")
```

Notice how:

* the `labs()` function is a separate function that we add on to the graph as a separate layer with the `+` operator
* inside the `labs()` function, we can supply an argument to the `title` parameter to change the title, and the `y` (and `x`) arguments to change the labels on those axes.

<!-- Book version -->

TODO: book exercises

<!-- End book -->

<!-- Tutorial version -->

<p class="comment3">
Using the `labs()` function, add a title and y-axis label to the boxplot of `price` that you created earlier (we will leave the x-axis label as its default, since that is adequate for this graph.)
</p>

```{r boxplot_labels_exercise, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Fill in the blanks:
diamonds %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = "price", y = price)) +
  _____(_____ = "A boxplot of price", _____ = "price ($)")
```

```{r boxplot_labels_exercise-solution, eval=is_tutorial, echo=FALSE}
# solution code
diamonds %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = "price", y = price)) +
  labs(title = "A boxplot of price", y = "price ($)")
```

```{r boxplot_labels_exercise-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->
