The `ggplot2` package gets its name because it's based on a concept called the *Grammar of Graphics* (hence **gg**plot...). This is a fascinating topic that we will learn more about later in this book, but the key idea is that **all data visualizations can be described using a common set of terms and ideas** (i.e. a "grammar"). The terms of this grammar correspond to the functions and their parameters in the `ggplot2` package. 

One of the most important parts of this grammar are the layers that we can add to a graph.

For example, the `ggplot()` function by itself just creates a blank canvas, i.e. the base layer of the graph:

```{r, echo=TRUE, warning=FALSE}
ggplot(data = starwars)
```

We can then add layers (created by other functions) by combining them with the `+` operator.

To get some kind of graph layer we need to take the __data__ and __map__ it to a __geometric__ shape.

For example, if we wanted to map the `height` column of the `starwars` dataset to the geometric shape of a histogram's bars, we could use this code:

`geom_histogram(mapping = aes(x = height))`

A few things to note:

* The `geom_histogram()` function is an example of what we call a *geom function* (each geom function specifies a different type of geometric shape).
* Geom functions have a parameter called `mapping` which tells R how to convert a column of data into a layer.
* The argument that we pass to the `mapping` parameter is a function called `aes()` (this is short for "aesthetic"). Inside the `aes()` function we need to specify any parts of the histogram's appearance that are determined by a column in the dataset.
  * In our example the x-axis of the histogram needs to show the `height` variable.

We then add this layer to the canvas created by `ggplot()` with the addition operator, `+`.

And so our final code looks like this:

```{r, warning=FALSE, message=FALSE, fig.show='hide'}
ggplot(data = starwars) +
  geom_histogram(mapping = aes(x = height))
```
