### Video overview

This video reviews the inference material covered in the first module.

<iframe width="560" height="315" src="https://www.youtube.com/embed/pVAxez7pA1M?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1uQMSTvCp1sWsH2OY1nLfUjpXaHKgRK6s/view?usp=sharing)


### Test your understanding

<!--
Other questions to add:
* What is observed statistic
* What is the null distribution
* interpreting the p-values meaning -> the probability of coming from the null
* alpha
-->

```{r response_vs_explanatory-quiz, eval=is_tutorial}
question(
  "You give sick patients either a drug or a placebo, and you measure whether they die or survive. This is stored in 2 variables, `treatment` and `outcome`, which are both categorical. What are your explanatory and response variables?",
  answer("response = `treatment`, explanatory = `outcome`"),
  answer("response = `outcome`, explanatory = `treatment`", correct = TRUE),
  answer('response = `"survived"`, explanatory = `"drug"`'),
  answer('response = `"survived"`, explanatory = `"placebo"`'),
  allow_retry = TRUE
)
```

```{r simulations_p_value-quiz, eval=is_tutorial}
question(
  "In a particular hypothesis test, we generate 10,000 permutations of our original data. 200 of them have a test statistic more extreme than our observed statistic. What is the p-value?",
  answer("0"),
  answer("0.01"),
  answer("0.02", correct = TRUE, message="The p-value is the probability that a value as or more extreme than the observed statistic came from the null distribution, which equivalent to the fraction of simulations more extreme than the observed statistic. This is 200/10,000 = 0.02"),
  answer("0.05"),
  answer("0.1"),
  answer("0.2"),
  answer("0.5"),
  answer("1"),
  answer("2"),
  answer("5"),
  answer("50"),
  allow_retry = TRUE
)
```



### Supplementary reading

These two chapters provide a good review of the original statistical inference material that we covered in the first inference module. You should review them before starting this week's assignment:

* Statistical Inference via Data Science by Chester Ismay and Albert Kim (available online at: https://moderndive.com/)
  * Chapter 7: Sampling - https://moderndive.com/7-sampling.html
  * Chapter 9: Hypothesis testing - https://moderndive.com/9-hypothesis-testing.html
  
On a lighter note, you can get an animated explanation of a hypothesis test with alpacas here: https://www.jwilber.me/permutationtest/

### Code review

In the first inference module, we used the `infer` package to simulate the null distribution.

We ran 10,000 random *permutations* to shuffle up the values of the response and explanatory variables. This random shuffling ensures that we simulate a distribution where there is no relationship between the two variables.

```r
<DATAFRAME> %>%
  specify(<RESPONSE> ~ <EXPLANATORY>) %>%
  hypothesize(null = "independence") %>%
  generate(10000, type = "permute") %>%
  calculate(stat = "...", order = c(..., ...))
```

* The `specify` function defines the response and explanatory variables.
* The `hypothesize` function declares what the null hypothesis is between those variables. `"independence"` means that there is no relationship between them.
* The `generate` function is where the simulations are actually run: we specify how many, and how to shuffle the data (in this case, using permutations).
* The `calculate` functions takes the simulations and calculates a *test statistic*: i.e. a single value for each one (e.g. the difference in proportions, or the difference in means).

<!--
### Test how much you remember!
-->
