The residuals of a linear model are arranged above and below the line. As we saw in the previous section, this arrangement should look similar to a normal distribution, i.e.

* most residuals will be small (close to the line)
* a few residuals will be large

The linear model also assumes that this normal distribution is the same in all parts of the model (i.e. at both high and low value of $x$).

For example, in this model, we can clearly see that the variation in the residuals *is not constant* - this is a violation of this third assumption of the linear model:

```{r}
set.seed(46)
x <- 1:100 + rnorm(100)
y <- x + 100:1 * rnorm(100)
ggplot(mapping = aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Decreasing variation in points")
```

In the graph above, the points on the right lie much closer to the line: i.e. the residuals will be smaller.

For simple linear models like this (with only a single explanatory variable $x$), we can get a reasonable idea of the variability of the residuals just by looking at a scatter plot of the model, as above.

For comparison, here is the model you fit earlier:

```{r}
sim_df %>%
  ggplot(mapping = aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


```{r assumption_3_quiz, eval=is_tutorial}
question(
  "Does your model (as shown above) satisfy the linear model's assumption of constant variability of residuals?",
  answer("Yes, the model satisfies this assumption.", correct = TRUE),
  answer("No, the model violates this condition."),
  allow_retry = TRUE
)
```
