### Create a multiple regression model

<!-- 
TODO
- switch to a smaller dataset that does not take so long to generate graphs
  - later note: does this make sense now that has been readded back into the initial modeling chapter? Ideally this chapter would share a unified dataset.
-->

Let's create a linear model from the `diamonds` dataset.

The `diamonds` dataset is included as part of the `ggplot2` package and contains observations about the properties of thousands of diamonds. We are going to use three variables from the dataset:

* the `price` variable will be our response variable (i.e. y)
* the `carat` and `depth` variables will be our explanatory variables

<!-- Book version -->

<!-- TODO: Add exercises -->

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=FALSE, results = 'asis', eval=is_tutorial}
cat('<p class="comment3">
Fill in the blanks to create this linear model.
</p>')
```

```{r kovzhddvrmedbxrc, exercise = TRUE, eval=is_tutorial, echo=FALSE}
diamonds_model <- lm(_____ ~ _______ + _______, data = _______)
```

```{r kovzhddvrmedbxrc-solution, eval=is_tutorial, echo=FALSE}
diamonds_model <- lm(price ~ carat + depth, data = diamonds)
```

```{r kovzhddvrmedbxrc-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```


<!-- End tutorial -->