### `lm()` with multiple explanatory variables

Recall that the first argument of the `lm` function was a formula. We create an R formula by:

* removing the constant coefficients
* replacing the $=$ sign with `~` (~ means "is related to")

For example,

$y = m x + c$

becomes

`y ~ x`

The same principle applies to multiple explanatory variables:


$y = m_1 x_1 + m_2 x_2 + c$

becomes

`y ~ x1 + x2`

<p class="comment3">
For each of the following questions, identify the correct R formula (or original linear model equation). Some questions assume that the data is in a data frame called `df`.
</p>

<!--
TODO:
- include feedback on these question
-->

```{r formula1_quiz, eval=is_tutorial}
question(
  "What is the correct R formula for this equation: ${Price} = m_1 \\times {Bedrooms} + m_2 \\times {Sqft} + c$ ?",
  answer("`y ~ x1 + x2`"),
  answer("`Bedrooms + Sqft ~ Price`"),
  answer("`Price + Bedrooms + Sqft`"),
  answer("`Price ~ Bedrooms + Sqft`", correct = TRUE),
  allow_retry = TRUE
)
```

```{r formula2_quiz, eval=is_tutorial}
question(
  "What is the equation behind this R formula: `cost ~ labor + goods` ?",
  answer("$y = m_1 x_1 + m_2 x_2 + c$"),
  answer("${cost} = m_1 \\times {labor} + m_2 \\times {goods} + c$", correct = TRUE),
  answer("$y = m_1 \\times {labor} + m_2 \\times {goods} + c$"),
  answer("`Cost ~ Labor + Goods`"),
  allow_retry = TRUE
)
```

```{r formula3_quiz, eval=is_tutorial}
question(
  "How would you create this linear model in R? ${Price} = m_1 \\times {Bedrooms} + m_2 \\times {Sqft} + c$",
  answer("`lm(Price ~ Bedrooms + Sqft, data = df)`", correct = TRUE),
  answer("`lm(Price ~ Bedrooms + Sqft)`"),
  answer("`df %>% lm(Price ~ Bedrooms + Sqft)`"),
  answer("`Price ~ Bedrooms + Sqft`"),
  allow_retry = TRUE
)
```

<!--
TODO
- actual code exercises on this
-->