We usually want to report the slope and intercept of a linear model. The slope is especially useful to report, because it indicates how much $y$ changes as $x$ changes.

* If we have a slope of 1, then this means that as $x$ increases, $y$ increases by the same amount (i.e. a ratio of 1).
* If we have a slope of 3, then as $x$ increases, $y$ increases by three times that value (e.g. if $x$ increases by 10, $y$ increases by $3 \times 10 = 30$).
* A slope of $0$ means that $x$ and $y$ are uncorrelated, i.e. as $x$ increases, the value of $y$ is not affected (on average). (We can still have variation in the $y$ variable - it's just that such variation seems to be entirely unconnected to changes in the $x$ variable).

<!--
TODO: graphs showing different slopes, corresponding to the three bullet point examples above, and include fig references back to this graph.
-->

The intercept is the point at which the line of our linear model intersects the $y$-axis. This is the value of the response variable when the explanatory variable is 0.

We can report the model's coefficients with the `tidy()` function from the `broom` package. The syntax is like this:

```r
tidy(your_model)
```

or equivalently

```r
your_model %>%
  tidy()
```

if you prefer to pipe the linear model in (the `your_model` variable above refers to the output of the `lm()` function).

For example, here is the table of coefficients for the model we were trying to create in the previous section:

```{r}
sim_model <- lm(y ~ x, data = sim_df)

sim_model %>%
  tidy()
```

The important column here (for our current purposes) is the *estimate* column. This tells us that the `lm()` function estimates the intercept to be `r coef(sim_model)["(Intercept)"]` and the slope to be `r coef(sim_model)["x"]`.

Note that although the intercept row is called `(Intercept)`, the slope row is named after the explanatory variable, `x`. As we shall see in the next section, we can have more than one explanatory variable in a single model, and this helps us keep track of the slope associated with each one.

<!-- TODO: show an example of creating the `tidy()` output table, and then how to interpret the numbers in it (since the column names are kind of weird).
-->


<!-- Book version -->

<!--
TODO: add book exercises
```{r, eval=is_book, echo=FALSE, results = 'asis'}
cat('
:::{.exercise}
Exercises TBD
:::
')
```
-->

<!-- End book -->

<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=FALSE, results = 'asis', eval=is_tutorial}
cat('<p class="comment3">
Replace the blank with the correct function to report the coefficients.
</p>')
```

```{r create_model, include=FALSE, eval=is_tutorial, echo=FALSE}
sim_model <- lm(y ~ x, data = sim_df)
```

```{r hojasacuvjabzyjp, exercise = TRUE, exercise.setup="create_model", eval=is_tutorial, echo=FALSE}
sim_model %>%
  _______()
```

```{r hojasacuvjabzyjp-solution, eval=is_tutorial, echo=FALSE}
sim_model %>%
  tidy()
```

```{r hojasacuvjabzyjp-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

```{r lm_slope_quiz, eval=is_tutorial, echo=FALSE}
question(
  "What is the slope of the linear model?",
  answer("1.27"),
  answer("1.75", correct = TRUE),
  allow_retry = TRUE
)
```

```{r lm_intercept_quiz, eval=is_tutorial, echo=FALSE}
question(
  "What is the intercept of the linear model?",
  answer("1.27", correct = TRUE),
  answer("1.75"),
  allow_retry = TRUE
)
```

<!-- End tutorial -->
