Let's create your first linear model in R, using a simulated[^define-simulated] dataset called `sim_df`.

<!-- TODO: convert to a real dataset -->

[^define-simulated]: A simulated dataset contains made-up values, usually generated randomly by a computer, rather than observations collected from real life.

This is what the first few rows of the simulated dataset stored in the `sim_df` variable look like:

```{r}
#| echo: false
sim_df %>% 
  head() %>%
  kableExtra::kable()
```

As you can see, there are two variables, $x$ and $y$. Here's what they look like:

```{r}
sim_df %>%
  ggplot() +
  geom_point(aes(x,y))
```

This looks like a pretty good candidate for a linear model, so let's create one.

We'll do that with the `lm()` function. The general syntax of this function is:

```r
lm(response ~ explanatory, data = _______)
```

* `response` and `explanatory` are the column names of those two variables. 
* The `data` parameter expects the dataframe containing those columns.

<div class="comment2">
What's going on with `response ~ explanatory`? This is what's know as an R *formula*.

The mathematical formula for a linear model is $y=ax+b$. In R, we would write this as `y ~ x`: i.e. we replace $=$ with `~`, and we only include the actual variables. From this, R infers that we want to estimate the model's coefficients (the slope, $a$, and the intercept, $b$).
</div>


<!-- Book version -->

```{r, eval=is_webbook, results='asis'}
#| echo: false
q_slope <- fitb("a")
cat(
  "<p>Exercise: In the linear model formula $y = ax + b$,</p>",
  "* The slope is represented by", q_slope
)
```
<!--
TODO: add book exercises
```{r, eval=is_book, echo=FALSE, results = 'asis'}
cat('
:::{.exercise}
Exercises TBD
:::
')
```
-->

<!-- End book -->


<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=FALSE, results = 'asis', eval=is_tutorial}
cat('<p class="comment3">
Fill in the blanks in the `lm()` function with the two variables, `y` and `x`, from our simulated dataset as well as the name of the dataframe, `sim_df`.
</p>')
```

```{r hclgfwuhnrvyquge, exercise = TRUE, warning=FALSE, message=FALSE, eval=is_tutorial, echo=FALSE}
sim_model <- lm(_____ ~ _______, data = _____)
```

```{r hclgfwuhnrvyquge-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"Make sure to get the variables in the right order!"
```

```{r hclgfwuhnrvyquge-solution, eval=is_tutorial, echo=FALSE}
sim_model <- lm(y ~ x, data = sim_df)
```

```{r hclgfwuhnrvyquge-check, warning=FALSE, message=FALSE, eval=is_tutorial, echo=FALSE}
grade_code()
```

<!-- End tutorial -->

<div class="comment3">
**Note**: Unfortunately the `lm()` function is a part of the core R programming language that predates the `tidyverse` packages. Unlike the tidyverse functions, the first argument of `lm()` is the formula, not the data frame. Therefore, you cannot pipe a data frame to `lm()`, because pipe inserts the data frame into the first argument position (which will cause an error).
</div>