### Linearity

By definition, a linear model's predictions (for the response variable) will all fall on a flat line/plane/hyperplane. Therefore, if the observed (i.e. true) values of the response variable are linearly related to the explanatory variables, they will also be linearly related to the predicted response variables.

In other words, no matter how many dimensions in our original model, we can always create a 2D plot to examine the assumption of linearity! We just have to create a *scatter plot* of the observed response variable vs the predicted response variable.

<!--
TODO
- need to explain this more thoroughly, ideally with a picture
-->

We call this an *observed vs predicted* plot (observed on the *y* axis, predicted on the *x*). Because these values are essentially the same number (or would be if the model went through every data point exactly), we should expect these variables to fall on a line with a slope of 1 and an intercept of 0. We will add such a reference line to our plot for comparison.

Before we can do so, we need to calculate the predicted values. We can easily create a column of predicted values using the `add_predictions` function from the `broom` package, which adds a column called `pred` to a dataframe using a model (such as the `diamonds_model` that you created earlier with the `lm` function):

```{r}
diamonds_model <- lm(price ~ carat + depth, data = diamonds)
```

```{r, echo = TRUE}
diamonds_plus_preds <- diamonds %>%
  add_predictions(diamonds_model)
```


<div class="comment3">
Using the new `diamonds_plus_preds` data frame, fill in the *3* blanks to create an *observed vs predicted* plot by:

* plotting the true (observed) response variable, `price`, on the y axis, and the model's predictions, `pred`, on the x axis
* adding a reference line using the `geom_abline` function with a slope of 1 and an intercept of 0.
</div>

```{r add_diamond_predictions, eval=is_tutorial}
diamonds_model <- lm(price ~ carat + depth, data = diamonds)
diamonds_plus_preds <- diamonds %>%
  add_predictions(diamonds_model)
```

```{r xwquzrffbtcvubwf, exercise = TRUE, exercise.setup = "add_diamond_predictions", eval=is_tutorial}
diamonds_plus_preds %>%
  ggplot() +
  geom_point(mapping = aes(x = _____, y = _____)) +
  geom_abline(slope = 1, intercept = ___, color = "red")
```

```{r xwquzrffbtcvubwf-solution, eval=is_tutorial}
diamonds_plus_preds %>%
  ggplot() +
  geom_point(mapping = aes(x = pred, y = price)) +
  geom_abline(slope = 1, intercept = 0, color = "red")
```

```{r xwquzrffbtcvubwf-check, eval=is_tutorial}
# check code
gradethis::grade_code()
```

We can see from this plot that while there is definitely a relationship, it is slightly non-linear. This is a violation of the linearity assumption.
