So we've reported the slope and intercept: however, these don't tell us about how well the model fits the data, i.e. how close the points are to the model's line.

For example, you could have a steep slope (suggesting that a change in $x$ is correlated with a big change in $y$), but the points could be scattered a long way from the line (i.e. large residuals). For example, consider these two graphs:

```{r}
set.seed(42)
x1 <- 1:10 + rnorm(10, 0, 0.5) 
x2 <- 1:10 + rnorm(10, 0, 0.5)
x <- c(x1, x2)
y <- c(x1 + rnorm(10, 0, 3), 2*x2 + rnorm(10, 0, 0.5))
series <- c(rep("Low slope, large scatter", 10), rep("Steep slope, low scatter", 10))
tibble(x, y, series) %>%
  ggplot(aes(x,y)) +
  geom_point() +
  geom_smooth(method="lm", se = FALSE) +
  facet_wrap(~ series)
```

We can use a statistic called $R^2$ (R-squared). You can think of $R^2$ as *the % of the variation in $y$ that is explained by the explanatory variable, $x$*.

$R^2$ varies between 0 and 1. 1 means that the model explains all the variation: i.e. the points fall exactly on the line of best fit! As the points get further from the line, $R^2$ will decrease.

We can report the $R^2$ by piping a model to the `glance()` function.

```r
your_model %>%
  glance()
```

<p class="comment3">
Fill in the blank with the correct function to report the $R^2$ value of your model.
</p>

```{r pejquogcchrafyyt, exercise = TRUE, exercise.setup="create_model", eval=is_tutorial}
sim_model %>%
  _______()
```


```{r pejquogcchrafyyt-solution, exercise.setup="create_model", eval=is_tutorial}
sim_model %>%
  glance()
```

```{r pejquogcchrafyyt-check, eval=is_tutorial}
# check code
gradethis::grade_code()
```

You'll note that `glance` reports a whole bunch of statistics. We only care about the first column, `r.squared`, and can ignore the rest.

```{r rsquared_quiz, eval=is_tutorial}
question(
  "What is the R-squared value of your model?",
  answer("0.935", correct=TRUE),
  answer("0.926"),
  answer("1.52"),
  answer("114"),
  allow_retry = TRUE
)
```
