So far we have looked at pieces of data by themselves:

```{r, echo=TRUE}
a <- 1
b <- 2
c <- 3
print(a)
print(b)
print(c)
```

But what about if we want to combine multiple pieces of data together?

R includes several types of *container* that can hold multiple pieces of data. We can then refer to that container by a single variable. For example, instead of the three variables above, we can create a **vector** that holds all three values. We create a vector with `c(...)`, putting the objects we want to combine inside the parentheses (and separated by commas):

```{r, echo=TRUE}
c(1,2,3)
```

*All the data in a vector must be the same type of data*. For example, a vector could contain all numbers, or all characters, but not a mix of the two.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Create a vector holding 3 character strings (in this order): "This", "is a", "vector!"
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Create a vector holding 3 character strings (in this order): "This", "is a", "vector!"
</p>')
```

```{r create_a_vector, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="create_a_vector">
<p class="comment">
**Hint:** No need to save this list to a variable...
</p>
</div>')
```

```{r create_a_vector-solution, eval=is_tutorial}
c("This", "is a", "vector!")
```

```{r create_a_vector-check, eval=is_tutorial, echo=F}
grade_code(":)")
```

<!-- End tutorial -->

> You might be wondering what the numbers in square brackets at the start of each line in the output mean? E.g. `[1]`
>
> These tell us where abouts in the vector we are. The number indicates the position in the vector of the first element displayed on that line.
>
> For example, the `[1]` at the start of the line (before "Introduction") shows that "Introduction" is the first element in this vector.

#### Operations on vectors

We can use operators on more complicated data structures just as we did on the simpler data types. For example, we can add 2 vectors together:

```{r, echo=TRUE}
v1 <- c(1,2,3)
v2 <- c(4,5,6)
v1 + v2
```
As you can see, the individual elements are added together.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
What happens if you add two vectors of different lengths? For example, run this code and see what happens:
:::
')
```

```{r, eval=is_book, results = FALSE, warning=FALSE}
v3 <- c(10, 20, 30, 40, 50)
v4 <- c(1, 2)
v3 + v4
```

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
What happens when you add two `v3` and `v4`?

Firstly, we get a warning `"longer object length is not a multiple of shorter object length"` because `v3` is longer than `v4`. However, a warning doesn\'t stop the code running - it merely tells us that something unexpected might be happening. 

In this particular case, R will do something called *recycling* which repeats the shorter vector over and over until it is the same length as the longer vector. I.e. `v4` will be repeated 2.5 times to become `(1,2,1,2,1)` before adding it to `v3`.

R warns you that this is happening because this may not be what you wanted, especially if you hadn\'t realized that the vecotrs were different lengths.
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
What happens if you add two vectors of different lengths?
</p>')
```

```{r vector_broadcasting, eval=is_tutorial, exercise=TRUE}
v3 <- c(10, 20, 30, 40, 50)
v4 <- c(1, 2)
v3 + v4
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="vector_broadcasting-hint">
<p class="comment">
**Hint:** Try changing the length of the shorter vector.
</p>
</div>')
```

```{r vector_broadcasting_quiz, eval=is_tutorial, echo=FALSE}
question("What happens when you add two vectors of different lengths?",
  answer("The 2 elements of v3 are added to the first 2 elements of v4, and the rest of v4 is unchanged."),
  answer("v4 is repeated to match the length of v3 before they are added.", correct=TRUE, message="This is called 'recycling', and is not what you might expect. It can cause bugs in R programs if the programmer (i.e. you) thinks that two vectors are the same length when they are not."),
  answer("R guesses values for the remaining values of v3 so that the two vectors are the same length."),
  allow_retry = TRUE
)
```

<!-- End tutorial -->
