### *If*-statements

There are also methods of **control flow** for choosing whether to skip a line entirely. We can use an `if` statement to run one or more lines of code based on the result of a Boolean condition:

```r
if(some_boolean_condition_is_TRUE){
  run the code inside these curly brackets
}
```

* A Boolean condition is an expression that results in a Boolean value (`TRUE` or `FALSE`). If the expression evaluates to `TRUE`, then the code following the *if* statements inside the curly brackets `{...}` is run.
* You are already familiar with these conditions, as you have used them in the `filter()` function to indicate which rows to pick from a dataframe.

<!-- Book version -->

<!-- TODO 
-->

<!-- End book -->

<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Change the value of the `x` variable in the first line so that the code in the if statement is run.
</p>
')
```

```{r qzxwzhahunefufkc, exercise = TRUE, echo=FALSE, eval=is_tutorial}
x <- "ninety-nine"

if(x == 99){
  print("You did it!")
}

print("This will always run.")
```
<!--
```{r qzxwzhahunefufkc-hint-1}
# hint text
""
```

```{r qzxwzhahunefufkc-hint-2}
# hint text
""
```
-->

```{r qzxwzhahunefufkc-solution, echo=FALSE, eval=is_tutorial}
x <- 99

if(x == 99){
  print("You did it!")
}

print("This will always run.")
```

```{r qzxwzhahunefufkc-check, echo=FALSE, eval=is_tutorial}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->


### *If/else*-statements

Often you may want to do one thing if a condition is TRUE, and something else if it's not. In this case you can put an `else{...}` block after the `if{...}` block:

```r
if(some_boolean_condition_is_TRUE){
  if TRUE, run the code inside these curly brackets
} else {
  otherwise, run this instead
}
```

<!-- Book version -->

<!-- TODO 
-->

<!-- End book -->

<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Change the Boolean value in the `if` parentheses so that the else block is executed instead.
</p>
')
```

```{r dxkfdawzkzncnspv, exercise = TRUE, echo=FALSE, eval=is_tutorial}
if(TRUE){
  print("the condition was TRUE")
} else{
  print("the expression was FALSE")
}
```

```{r dxkfdawzkzncnspv-hint-1, echo=FALSE, eval=is_tutorial}
# hint text
""
```

```{r dxkfdawzkzncnspv-hint-2, echo=FALSE, eval=is_tutorial}
# hint text
""
```

```{r dxkfdawzkzncnspv-solution, echo=FALSE, eval=is_tutorial}
if(FALSE){
  print("the condition was TRUE")
} else{
  print("the expression was FALSE")
}
```

```{r dxkfdawzkzncnspv-check, echo=FALSE, eval=is_tutorial}
# check code
gradethis::grade_code()
```

In general you won't want to put just `TRUE` or `FALSE` in the parentheses in real world code, because it will always execute one of the blocks or the other - however it is a valid Boolean expression!

<!--
multiple if/else

### The `if_else()` function

Sometimes we want to 

if_else inside mutate

case_when inside mutate

True and False as 0 and 1

-->