### Vectors

You can create a vector with the `c()` function. You can also create a vector of numbers using the `:` operator. For example, these two lines both create the same vector of the values 1 through 5:

```{r, echo=TRUE}
c(1,2,3,4,5)
```

```{r, echo=TRUE}
1:5
```

<!-- Book version -->

<!-- TODO 
example code from first R programming chapter:

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Run these two lines of code in the RStudio Console and see what result you get.

`a <- 2 + 2`

`5 * a`

Then take a look at the *Environment* tab in the top-right pane of RStudio. 
Do you see a __variable__ called `a`? 
Does it hold the value calculated in the first line of code or the second?
:::
")
```
-->

<!-- End book -->

<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Create a vector of all the integer numbers from 2 to 250.
</p>
')
```

```{r aiwkmkcqayvrufhf, eval=is_tutorial, echo=F, exercise=TRUE}

```

```{r aiwkmkcqayvrufhf-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"Try using the `:` operator"
```

```{r aiwkmkcqayvrufhf-solution, eval=is_tutorial, echo=FALSE}
2:250
```

```{r aiwkmkcqayvrufhf-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```


### Other vector tips

There are several other useful things to remember about vectors:

* You can combine two vectors into a single vector with the `c()` function:

  ```{r, echo=TRUE}
  x <- 1:3
  y <- 4:6
    
  c(y, x)
  ```

* You can *extract* a single value from a vector by **indexing** into the vector with square brackets `[...]` and the position of item that you want:

  ```{r, echo=TRUE}
  x <- 101:110
    
  x[5]
  ```

* You can change any value in a vector with square brackets and the assignment operator:

  ```{r, echo=TRUE}
  x <- 1:5
  x[2] <- 42
  
  x
  ```

* You can even extend a vector by one element by assigning to a position 1 higher than it's current length:

  ```{r, echo=TRUE}
  x <- 1:5
  x[6] <- 42
    
  x
  ```

<!-- TODO: 
Add exercises
-->