### Combining data with operators

Okay, so now we know about data.

But data by itself is not especially useful. It just sits there until you *do something* to it. In programming languages there are often many ways of doing things to data, but one of the simplest is **operators**.

Operators *operate* on data. The name may sound unusual, but you are already familiar with many operators, such as `+` and `-` for adding and subtracting numbers.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Try entering a number after the `>` in the Console (e.g. 1), then Enter, and see what happens.
:::

When you hit enter, the R interpreter reads in the line, evaluates it, and returns the answer. In this case, you entered 1, so the computer thinks 'Hey, it’s a 1! Wow, a one! The result of 1 is… drum roll, please… 1!' and returns the result of this **expression**, which is a one.

Cool! But not, I confess, particularly useful. Let’s fix that: next we'll add two numbers together. 

:::{.exercise}
At the prompt, enter two numbers separated by a plus sign, +

`> 1 + 1`

What do you get?
:::

(Note that I've left the Console's `>` prompt in the example code above, but I will leave it out in future.)
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Let's try using the addition operator, `+`, on some numbers to see what happens. Run this code chunk:
</p>
")
```

```{r operators_simple_addition, eval=is_tutorial, echo=F, exercise=TRUE}
1 + 1
```

```{r operators_simple_addition-check, eval=is_tutorial, echo=F}
grade_result(
  pass_if(~ identical(.result, 2), "Unsuprisingly, the + operator adds two numbers together.")
)
```

<!-- End tutorial version -->

Great! Let's move on and investigate operators in more depth...

### Operating on numbers

Heart surgeons operate on hearts, brain surgeons operate on brains. 
You will be operating on numbers... does that make you a data surgeon?

Here are some of the operators available to us in R:

| Operator  | Example  | Result  |
|:---------:|:--------:|:-------:|
| `+`       | `5 + 2`  | `7`     |
| `-`       | `5 - 2`  | `3`     |
| `*`       | `5 * 2`  | `10`    |
| `/`       | `5 / 2`  | `2.5`   |
| `^`       | `5 ^ 2`  | `25`    |
| `%%`      | `5 %% 2` | `1`     |
| `<`       | `5 < 2`  | `FALSE` |
| `>`       | `5 > 2`  | `TRUE`  |
| `==`      | `5 == 2` | `FALSE` |

Some of these might seem obvious, while others might be unfamiliar. 
Let's go through them all and figure out what they do.

### The `-` operator

<!-- Book version -->
```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
In the R Console, type `5 - 2` and hit enter to run the line of code.

You probably have a good idea of what `-` does, but try changing the numbers
just to make sure!
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `-` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_subtraction, exercise=TRUE, eval=is_tutorial}
5 - 2
```

```{r operators_subtraction_quiz, echo=FALSE, eval=is_tutorial}
question("What does the - operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another.", correct = TRUE),
  allow_retry = TRUE
)
```


### The `*` operator

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
In the R Console, type `3 * 2` and hit enter to run the line of code.

What does `*` do?
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `*` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_multiplication, eval=is_tutorial, exercise=TRUE}
3 * 2
```

```{r operators_multiplication_quiz, eval=is_tutorial, echo=FALSE}
question("What does the * operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another."),
  answer("Multiplies two numbers together.", correct = TRUE),
  answer("Divides one number by another."),
  allow_retry = TRUE
)
```


### The `/` operator

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
In the R Console, type `3 / 2` and hit enter to run the line of code.

What does `/` do? Just to be sure, try some other numbers.
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `/` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_division, eval=is_tutorial, exercise=TRUE}
3 / 2
```

```{r operators_division_quiz, eval=is_tutorial, echo=FALSE}
question("What does the / operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another."),
  answer("Multiplies two numbers together."),
  answer("Divides one number by another.", correct = TRUE),
  allow_retry = TRUE
)
```


### The `^` operator

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
In the R Console, type `3 ^ 2` and hit enter to run the line of code.

What does `^` do? Try some other numbers like `2 ^ 3` or `16 ^ 0.5`
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `^` operator does (play around with the numbers to see what happens):
</p>
")
```


```{r operators_exponent, eval=is_tutorial, exercise=TRUE}
4 ^ 2
```

```{r operators_exponent_quiz, eval=is_tutorial, echo=FALSE}
question("What does the ^ operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second).", correct = TRUE),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```


### The `%%` operator

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Next up, a slightly trickier one, type `3 %% 2` and hit enter to run the line of code.

What does `%%` do? You will probably have to try some other numbers to figure 
this one out.

If you have difficulty try also dividing the same numbers. E.g. try both 
`9 %% 4` and `9 / 4`.
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `%%` operator does (play around with the numbers to see what happens - for instance, try `4 %% 3` as well as `6 %% 3` and `3 %% 3`):
</p>
")
```



```{r operators_modulo, eval=is_tutorial, exercise=TRUE}
5 %% 3
```

```{r operators_modulo_quiz, eval=is_tutorial, echo=FALSE}
question("What does the %% operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another).", correct = TRUE),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```


### The `<` and `>` operators

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
So far we have operated on numbers and got numbers back. However these next 
operators will return a different type of data.

What do you get if you run `8 > 10`?

Can you change the `8` into a number so that this expression returns `TRUE`?

Then change the `>` to a `<` operator (i.e. reverse its direction). 
What is the result now?
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `<` operator does.

Do you get a number back (like the previous operators), or a different type of data?

Then change the `<` to a `>` operator (i.e. reverse its direction). What is the result now?
</p>
")
```

```{r operators_inequality, eval=is_tutorial, exercise=TRUE}
4 > 3
```

```{r operators_inequality_quiz, eval=is_tutorial, echo=FALSE}
question("What does the > operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another).", correct = TRUE),
  allow_retry = TRUE
)
```

> __Boolean operators__
> 
> In the previous section we learned about the Boolean data type which has only
> two possible values: `TRUE` or `FALSE`.
>
> Operators that always return a Boolean value are called Boolean operators.
> The *greater than* (`>`) and *less than* (`<`) operators are examples of
> Boolean operators.
>
> You can think of Boolean operators as asking a question that can only be
> be answered with either "true" or "false", such as "Is 8 greater than 10?"

### The `==` operators

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Let's try another Boolean operator. What do you get if you run `8 == 10`?
What about `8 == 8`?

What do you think the `==` operator does?
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `==` operator does.

Then try changing one of the numbers and rerunning the code.
</p>
")
```

```{r operators_equality, eval=is_tutorial, exercise=TRUE}
4 == 4
```

```{r operators_equality_quiz, eval=is_tutorial, echo=FALSE}
question("What does the == operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another).", correct = TRUE),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```


> __Combining comparisons__
>
> Sometimes we want to know if one datum is greater than **or** equal
> to another. You can use the Boolean operators `>=` for such a comparison, or
> `<=` to see if something is *less than or equal to* another.


### Which operator goes first?

Just like in normal math, we can do sums in R with multiple operators:

```r
3 + 5 / 5 * 3 ^ 2
```

In such a case, which operation do we do first?

Again, just like in regular math, some operations are always done before others. For example, all multiplication and division will be done before any addition or subtraction.

> **F.Y.I.**
>
> The order in which operators are calculated is known as *operator precedence*, and you can find the precedence of any operator here: https://stat.ethz.ch/R-manual/R-devel/library/base/html/Syntax.html

We can change the order of operations with parentheses: `(` and `)`. For example

```r
2 + 2 * 5 = 12
```

whereas

```r
(2 + 2) * 5 = 20
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Modify this R code

`3 + 5 / 5 * 2 ^ 2`

so that it performs the calculation $\frac{3 + 5}{(5 \times 2) ^ 2}$.

When correct, you should get the answer 0.08.
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Modify the expression below so that it performs this calculation $\frac{3 + 5}{(5 \times 2) ^ 2}$.
</p>
")
```

```{r operators_parentheses, eval=is_tutorial, exercise=TRUE}
3 + 5 / 5 * 2 ^ 2
```

```{r operators_parentheses-check, eval=is_tutorial}
grade_result(
  pass_if(~ identical(.result, 0.08), "Looks like you're a parenthesis master!")
)
```
