It is a long-standing tradition that your first program in any new programming language should simply display the message "Hello, World!".

In R we can use the `print()` *function* to display a piece of data (such as the text *Hello, World!*).

For example, we can run this line of code in R:

```{r, results=FALSE}
print("Hello, World!")
```

And R will return this result:

```{r, echo=FALSE}
print("Hello, World!")
```

::: {.info-box}
> **What's happening here?**
>
> We will learn more about the different parts of this code in this chapter,
> but briefly:
>
> * `print(...)` is a **function**: it takes some input ingredients (which we 
>   list between the brackets after the function's name), and does something 
>   with them.
> * `"Hello, World!"` is a string of characters (and other symbols, such as 
>   `!`). A **character string** is a type of data that computers can work with.
:::

<!-- Book code -->

```{r hello-world-exercise-book, echo=F, eval=is_book, results = 'asis'}
cat("::: {.exercise}
**Exercise: Try it yourself:** Open up RStudio and copy the line of code into the RStudio
Console. Then hit the <Enter> key on your keyboard to run the line. You should
get back the same output as above.

Congratulations! You just ran your first program in R.
:::")
```

<!-- Tutorial code -->

```{r hello-world-exercise-instructions-tutorial, echo=F, eval=is_tutorial, results = 'asis'}
cat('<p class="comment3">
Run the simple program in the code section below (click on *Run Code*). The output of the program will be displayed below the code section. Then edit the text inside the `print()` function to the correct text (*Hello, World!*), and try running the code again. Once it displays the correct statement, click the *Submit Answer* button.

(If a yellow box appears warning you that your answer is "invisible", don\'t worry. You can safely ignore this.)
</p>')
```

```{r hello-world-exercise-tutorial, eval=is_tutorial, exercise=TRUE, echo=F}
print("Hello, Student!")
```

```{r hello-world-exercise-tutorial-hint, echo=F, eval=is_tutorial, results = 'asis'}
cat('<div id="hello-world-hint">
<p class="comment">
**Hint:** Enter the correct text inside quotation marks.
</p>
</div>')
```

```{r hello-world-exercise-tutorial-solution, eval=is_tutorial, echo=F}
print("Hello, World!")
```

```{r hello-world-exercise-tutorial-check, eval=is_tutorial, echo=F}
gradethis::grade_code("You just ran your first program! Don't worry, things will get more interesting soon...")
```
