Perhaps, keen mathematician that you are, you want to calculate the length of the hypotenuse of a triangle. Dredging up memories of early math classes, you will doubtless recall Pythagoras’s theorem that the hypotenuse (the long side) of triangle is given by:

$c = \sqrt{a^2 + b^2}$

($c$ is the hypotenuse [long side] and $a$ and $b$ are the short sides.)

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Let’s say we have a triangle where the shorter sides (a & b) are 3 and 4 units long. Can you calculate the length of side c in R using just the operators from the first section?

*Hint #1:* The square root is equal to the 0.5 power of a number: `4 ^ 0.5 = 2`

*Hint #2:* Just like in regular math equations, R will calculate some operators before others. For example it will do all multiplications before any additions. However, just like in regular math, we can change the order of operations by wrapping parts of our calculation in parentheses: `(...)`
:::

Did you get the answer 5? Fantastic!
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Write one (or more) lines of code that calculates the hypotenuse of a triangle where the short sides have lengths 3 and 4.
</p>')
```

```{r hypotenuse_code, exercise=TRUE, eval=is_tutorial}

```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="hypotenuse_code-hint">
<p class="comment">
**Hint:** Remember that the square root is the same as raising a number to the power of $\\sqrt{x} = x^\\frac{1}{2}$, i.e. in R: `x^0.5`.
</p>
</div>')
```

```{r hypotenuse_code-check, eval=is_tutorial, echo=F}
grade_conditions(
  pass_if(~ .result == sqrt(10^2 + 20^2))
)
```

<!-- End tutorial -->

### Re-useable code = functions

What’s that? Another complaint? You have to write out this long expression every time you need the hypotenuse of a triangle? (No doubt this is a frequent chore in your day-to-day life.)

Again, there is a solution! R allows us to save pieces of code in variables.  Yes, you heard that right: variables don’t just have to store data, they can also store code!

These stored, reusable sections of code are called **functions**.

For example, you could create a function to calculate the sum of two numbers:

    adder <- function(number1, number2) {
        result <- number1 + number2
        return(result)
    }


Entering these 4 lines at the console prompt will be slow and error-prone, so let’s try something different.

Click on the "File" menu at the top of RStudio. Select "New File" and then "R Script". A blank editor window should appear in a new pane above the console.

Copy the adder function from the previous page into this empty script. Then press "Control + Alt + R" on your keyboard (simultaneously). This will run the contents of your script all at once.

If successful, you should see that `adder` appears in the Environment pane under a new section called *Functions*.

How do we use our adder function? Go back to the console, and type something like this:

> adder(3, 5)

If your function is working correctly you should get the result of the 2 numbers that you entered inside the braces.

Let's take another look at the adder function to understand what's going on:

    adder <- function(number1, number2) {
        result <- number1 + number2
        return(result)
    }

**Line 1:** The first line creates a new function with the `function` keyword and saves it to the name `adder` using the assignment operator `<-`, just as we did for variables. 

After `function` are a pair of parentheses. Inside these, we put a list of the parameters that the function can take, separated by commas. In this case, our adder function has two paramters (the numbers to add together). We are going to give these numbers the temporary names `number1` and `number2` (creative, I know). We will use these parameter names inside the function to refer to these two numbers.

We end the line with an opening curly bracket `{` to indicate that the code that follows is part of the function.

**Line 2:** This is the meat of our adder function. We add our two number paramters together and store them in a variable called `result`. Its important to note that `result` only exists inside the curly brackets of the adder function (i.e. it vanishes after the function has finished).

**Line 3:** Here we specify what the function is should return: in this case we want to return the `result` variable.

**Line 4:** We signal the end of the function with a closing curly bracket (matching the one from the end of line 1).

You might object (and not without reason) that our `adder` function is a very trivial example. Wouldn't it just be easier to use the `+` operator?

Yes, it would! So let's look at a more complicated function.

We can create a function to calculate the hypotenuse like this:

```{r, echo=TRUE}
hypotenuse <- function(a, b) {
  c <- (a^2 + b^2)^0.5
  return(c)
}
```

Then we can use this `hypotenuse` function as many times as we like. For example calculate the hypotenuse of a triangle with sides of length 3 and 4, we would run:

```{r, echo=TRUE}
hypotenuse(3, 4)
```


<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Use the `hypotenuse()` function to calculate the area of a triangle with sides of length 3 and 4.

*Hint:* Try changing the numbers inside the parentheses after `hypotenuse`.
:::

Did you get the answer 5? Fantastic!
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Use the `hypotenuse()` function to calculate the area of a triangle with sides of length 3 and 4.
</p>')
```

```{r hypotenuse_definition, eval=is_tutorial}
hypotenuse <- function(a, b) {
  c <- (a^2 + b^2)^0.5
  return(c)
}
```

```{r hypotenuse_function, exercise=TRUE, exercise.setup = "hypotenuse_definition", eval=is_tutorial}

```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="hypotenuse_function-hint">
<p class="comment">
**Hint:** Try changing the numbers inside the parentheses after `hypotenuse`
</p>
</div>')
```

```{r hypotenuse_function-check, eval=is_tutorial, echo=F}
grade_conditions(
  pass_if(~ .result == sqrt(3^2 + 4^2))
)
```

<!-- End tutorial -->

### How the `hypotenuse` function works

There are a few things to note about this code:

* We tell R that we are creating a reusable function using the `function` keyword.
* `function` is followed by parentheses `(...)` that contain **parameters**. Parameters are the names that we give to the input data to the function.
  * For example, above we created two parameters: `a` and `b`
  * You can have as many parameters as you want in a function, from zero on up. They must be separated by commas.
* The reusable code goes inside a pair of curly brackets `{...}`
  * We can now use the function's parameters in this code (e.g. `a` and `b`). Essentially we temporarily create new variables with the parameter names (but these are)
* At the end of the function we can return a particular result with `return(...)` - just replace the dots with a value or 
* We store the function in a name with the assignment operator `<-` (just like we did with variables)
* When we want to run the code, we write the function name followed by parentheses, with any arguments inside the parentheses (separated by commas)

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Replace the blanks to create a function to calculate the *area of a triangle* instead. Save this function as `triangle_area`.

_______ <- function(a, b) {
  area <- _______
  return(area)
}

*Hint:* The area of a triangle is $0.5 \times a \times b$.
:::
')
```

<!-- End book -->


<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Replace the blanks to create a function to calculate the *area of a triangle* instead. Save this function as `triangle_area`.
</p>')
```

```{r triangle_area_function, exercise=TRUE, eval=is_tutorial, echo=F}
_______ <- function(a, b) {
  area <- _______
  return(area)
}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="triangle_area_function-hint">
<p class="comment">
**Hint:** The area of a triangle is `0.5 * a * b`.
</p>
</div>')
```

```{r triangle_area_function-check, eval=is_tutorial, echo=F}
grade_conditions(
  pass_if(~ identical(.result(4,5), 10)),
  pass_if(~ identical(.result(10,20), 100))
)
```

<!-- End tutorial -->

<!--
TODO

### Create your own functions
or
## Additional extra exercises


Return a fixed value.

Calculate the area of a rectangle.

Passing variables, & making copies, modifying within function, & global variables

-->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("### The End!

That's all for this tutorial, folks!")
```
