### Combining different data into lists

TODO: section moved from first R chapter - needs rewriting

So far we have looked at pieces of data by themselves:

```{r, echo=TRUE}
a <- 1
b <- "Hello!"
c <- TRUE
print(a)
print(b)
print(c)
```

But what about if we want to combine multiple pieces of data together?

R includes several types of *container* that can hold multiple pieces of data. We can then refer to that container by a single variable. For example, instead of the three variables above, we can create a **list** object that holds all three values:

```{r, echo=TRUE}
l <- list(1, "Hello!", TRUE)
l
```

We create a list using `list()`. Every value in the list goes inside the parentheses, separated by commas.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Create a list holding 4 values (in this order): 10, "z", FALSE, -0.1*10
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Create a list holding 4 values (in this order): 10, "z", FALSE, -0.1*10
</p>')
```

```{r create_a_list, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="create_a_list-hint">
<p class="comment">
**Hint:** No need to save this list to a variable...
</p>
</div>')
```

```{r create_a_list-solution, eval=is_tutorial}
list(10, "z", FALSE, -0.1*10)
```

```{r create_a_list-check, eval=is_tutorial, echo=F}
grade_code(":)")
```

<!-- End book -->