<!-- TODO:
video overview of R variable creation
-->

So far we have learnt how to combine data to get different results.

We can do multiple separate calculations by putting each one on a separate line. When R reads your code, it treats everything on one line as a single expression that is separate from other lines:

```r
2 + 2
5 * 5
```

This program will have two separate outputs: 4 and 25

However, after these results are shown to us, they are thrown away! All that effort just discarded...

What if we wish to save the result of a calculation so that we can reuse it in a subsequent line?

In this case, we need to store the result in a **variable**.

<!-- Book version -->

<!-- TODO:
add a hint showing where the environment is in RStudio via a screenshot. Might be helpful to also add this screenshot to the tutorial version as well.
-->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Run these two lines of code in the RStudio Console and see what result you get.

`a <- 2 + 2`

`5 * a`

Then take a look at the *Environment* tab in the top-right pane of RStudio. 
Do you see a __variable__ called `a`? 
Does it hold the value calculated in the first line of code or the second?
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Run these two lines of code and see what result you get.
</p>
')
```

```{r operators_assignment, eval=is_tutorial, echo=F, exercise=TRUE}
a <- 2 + 2
5 * a
```

```{r operators_assignment_quiz, eval=is_tutorial, echo=FALSE}
question("What has just happened?",
  answer("The first line calculates 2 + 2 and the second line calculates 5 * 5."),
  answer("The value 2 was stored in the variable a, and then the value of a was used in the second line."),
  answer("The result of 2 + 2 was stored in the variable a, and then the value of a was multiplied by 5 in the second line.", correct = TRUE),
  allow_retry = TRUE
)
```


### The "result" of the assignment operator

We store the result of an expression in a variable using the **assignment operator**: `<-`

```r
variable_name <- value_to_be_stored
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Run these two lines of code in the RStudio Console and see what result you get.

`2 + 3`

`b <- 2 + 4`

If you take another look at the *Environment* tab in the top-right pane of RStudio, you should see another variable called `b`. What value does it hold?
:::
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
What is the *output* of running an expression with and without assigning it to a variable? Run the two code blocks below.
</p>
')
```


```{r sum_without_assignment, eval=is_tutorial, exercise=TRUE}
2 + 3
```

```{r sum_with_assignment, eval=is_tutorial, exercise=TRUE}
b <- 2 + 4
```

<!-- End tutorial version -->

The second line of code should not print out any output when it runs. 
This is because assigning the result of an expression to a variable has no "result" in itself. For example, in math $2+2$ is $4$, but the expression $b = 2 + 2$ does *not* return $4$ directly (but somebody somewhere is hopefully keeping track of the fact that $b$ is now equivalent to $4$).

If you want to see the data that is stored in a variable, you can put the name of the variable on a line by itself:

```r
some_variable
```

R will evaluate this line: it will ask itself "What is the *result* of `some_variable`", which is just whatever value is stored in that variable.

For example,

```{r}
c <- 3

c
```

The other implication of this is that if you calculate something in R and *do not* assign the result to a variable then it will be printed out ***and then forgotten***. 
So remember: if you calculate something important in R that you will need in the future, make sure that you store that result in a variable.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Type the name of one of the variables in your Environment tab in the RStudio Console (e.g. `a`), and hit Enter to run it.

Does this return the data that you think is stored in that variable?
:::
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r what_is_b_setup, eval=is_tutorial, echo=FALSE}
b <- 6
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Output the value of the variable `b` that you created above.
</p>
')
```


```{r what_is_b, eval=is_tutorial, exercise=TRUE, exercise.setup = "what_is_b_setup"}
```


```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="what_is_b-hint">
<p class="comment">
**Hint:** Try putting the variable `b` on a line by itself...
</p>
</div>
')
```

```{r what_is_b-solution, eval=is_tutorial, echo=FALSE}
b
```

```{r what_is_b-check, eval=is_tutorial, echo=F}
grade_code(":)")
```

<!-- End tutorial version -->

### Variables are... variable

Variables get their name because *their value can vary*. We have created the variable `b` that holds the value 6, but we can change the value of `b` and store a completely different value in it!

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Try assigning the value `7` to the variable `b` using the assignment operator `<-`.

Since we already created `b` in an earlier exercise, you should see that its value in the Environment tab updates.
:::
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Assign the value `7` to to the variable `b`
</p>
')
```

```{r b_is_7, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="b_is_7-hint">
<p class="comment">
**Hint:** Try using `<-` to assign a value to `b`
</p>
</div>
')
```

```{r b_is_7-solution, eval=is_tutorial}
b <- 7
```

```{r b_is_7-check, eval=is_tutorial, echo=F}
grade_result(
  pass_if(function(x) identical(x, 7), "This is a correct message.")
)
```

<!-- TODO: weird output needs fixing -->

<!-- End tutorial version-->

### When does assignment happen

\index{precendence}

`<-` is an operator, just like `+` or `*`. As such, it has a *precedence*: it will happen before some operators but after others.

However, it turns out that the `<-` precedence is [extremely low](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Syntax.html) - i.e. it will happen after the result of all the other operators on that line of code have been calculated.

So, when you write:

```r
a <- 2 + 4
```

...you are essentially doing this:

```r
a <- (2 + 4)
```

<!-- TODO: 
Add exercises e.g. remove parentheses from over parenthesized expression; split line of code over multiple lines.
-->


