### Boolean data revisted

At the start of this chapter we mentioned that there is a type of data in R, called Boolean data, that can have one of two values: `TRUE` or `FALSE`.

We can ask R questions that have a true or false answer, for example: "Does the variable `x` hold the number 3?" or "Is 10 greater than 9?"

We do this with *Boolean operators*:

| Operator  | Example  | Result  |
|:---------:|:--------:|:-------:|
| `<`       | `10 < 9` | `FALSE` |
| `>`       | `10 > 9` | `TRUE`  |
| `==`      | `x == 3` | `FALSE` |

For example:

```{r, include=TRUE, echo=TRUE}
10 < 9
```

Here R returns the value `FALSE` when it evaluates this expression, because 10 is obviously not less than 9.

> __Combining comparisons__
>
> Sometimes we want to know if one datum is greater than **or** equal
> to another. You can use the Boolean operators `>=` for such a comparison, or
> `<=` to see if something is *less than or equal to* another.

Just as with numeric data, we can store a Boolean value in a variable, e.g. `d <- FALSE` or `e <- 10 < 9`. (Remember that this assignment to a variable always happens last, after we have evaluated the expression on the right-hand side of the `<-` operator.)

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("TODO: 
put exercises into their own section ")
```


### The `<` and `>` operators

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("TODO: covert exercises to render functions")
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
**Exercise: Try it yourself:** What do you get if you run `10 < 9` in the RStudio Console?

Can you change one of the numbers so that this expression returns `TRUE`?

Then change the `>` to a `<` operator (i.e. reverse its direction). 
What is the result now?
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `<` operator does.

Do you get a number back (like the previous operators), or a different type of data?

Then change the `<` to a `>` operator (i.e. reverse its direction). What is the result now?
</p>
")
```

```{r operators_inequality, eval=is_tutorial, exercise=TRUE, echo=is_tutorial}
4 > 3
```

```{r operators_inequality_quiz, eval=is_tutorial, echo=FALSE}
question("What does the > operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another).", correct = TRUE),
  allow_retry = TRUE
)
```

### The `==` operators

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Let's try another Boolean operator. What do you get if you run `8 == 10`?
What about `8 == 8`?

What do you think the `==` operator does?
:::
")
```

<!-- End book

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `==` operator does.

Then try changing one of the numbers and rerunning the code.
</p>
")
```

```{r operators_equality, eval=is_tutorial, exercise=TRUE, echo=is_tutorial}
4 == 4
```

```{r operators_equality_quiz, eval=is_tutorial, echo=FALSE}
question("What does the == operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another).", correct = TRUE),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```

<!-- End tutorial -->


> __Combining comparisons__
>
> Sometimes we want to know if one datum is greater than **or** equal
> to another. You can use the Boolean operators `>=` for such a comparison, or
> `<=` to see if something is *less than or equal to* another.




<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Assign the value `TRUE` to a variable called `d`.
:::
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Assign the value `TRUE` to a variable called `d`.
</p>
')
```

```{r d_is_true, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="d_is_true-hint">
<p class="comment">
**Hint:** Try using `<-` to assign a Boolean value to a variable called `d`
</p>
</div>
')
```

```{r d_is_true-solution, eval=is_tutorial}
 d <- TRUE
```

```{r d_is_true-check, eval=is_tutorial, echo=F}
grade_code(":)")
```

<!-- End tutorial -->