### *For* loops

The most common type of loop in R is the **for loop**.

The idea behind the for loop is that "*for* every *thing* in a *group of things*, run some code".

In R code, we would write this as:

```r
for(thing in many_things){
  some code to run
}
```

For example, you could loop through a vector of numbers, and print out each number:

```{r wyfjiaqcdwwcrasz, exercise = is_tutorial}
for(x in 1:10){
  print(x)
}
```

<!--
```{r wyfjiaqcdwwcrasz-hint-1}
# hint text
""
```

```{r wyfjiaqcdwwcrasz-hint-2}
# hint text
""
```

```{r wyfjiaqcdwwcrasz-solution}
for(x in 1:10){
  print(x)
}
```

```{r wyfjiaqcdwwcrasz-check}
# check code
gradethis::grade_code()
```
-->

Note that on each **iteration** of the loop we update the variable `x` (or whatever you choose to call it) and that variable can be used in the code inside the for loop's code block `{...}`

<p class="comment3">
Fill in the blank in the *if* statement's condition so that only *even* numbers are printed out. (Remember that an even number divided by 2 will have have a remainder of 0.)
</p>

```{r yrmkqxvwpawfdzeo, exercise = TRUE, eval=is_tutorial}
for(x in 1:10){
  if(_____){
    print(x)
  }
}
```

```{r yrmkqxvwpawfdzeo-hint-1, eval=is_tutorial}
# hint text
""
```

```{r yrmkqxvwpawfdzeo-hint-2, eval=is_tutorial}
# hint text
""
```

```{r yrmkqxvwpawfdzeo-solution, eval=is_tutorial}
# solution code
for(x in 1:10){
  if(x %% 2 == 0){
    print(x)
  }
}
```

```{r yrmkqxvwpawfdzeo-check, eval=is_tutorial}
# check code
gradethis::grade_code()
```
