The central component of everything we will be doing in R this semester is **data**. Even non-data science programs revolve around data. 

At a very basic level, a computer is just a fancy calculator that adds and subtracts numbers. Even things like words and pictures are stored inside a computer as numbers.

However, we often want to work work with data that is not numbers. For example, in the last section we were able to get R to print out the sentence *Hello, World!* To your computer that was just numbers flowing down wires as electrical signals. But the R programming language took care of converting our instruction into something your computer could understand.

This is the magic of programming languages! They allow us to write commands in (relatively) human-readable instructions, and then take care of translating that into the very unreadable numbers that computers work with.

R allows us to work with several "higher-level" types of data. These **data types** include:

* the **character** data type holds text (i.e. letters, symbols, and the characters that represent numbers). We need to put the text inside quotation marks so that R knows where the text starts and ends: `"this is character data"`.

  * Note: in other programming languages this datatype is sometimes known as a "character string" or just a "string".

* the **numeric** data type holds numbers such as `0` or `12.5` (i.e. both integers (whole numbers) and decimals). Unlike text, numbers are written *without* quotation marks around them.

* the **Boolean** data type holds a value that is either `TRUE` or `FALSE`. (This is sometimes also referred to as the logical data type.)

<!-- Book version -->

```{r numeric-data-book-exercise, eval=is_book, echo=F, results='asis'}
cat('::: {.exercise}
**Exercise:** Type the number 2 into the RStudio console. Hit the <Enter> key to run this line of code. What do you get back?
:::')
```

<!-- End book version -->

<!-- Tutorial version -->

```{r character_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
gradethis::question('What data type is `"CDS 101"`?',
  answer("Numeric."),
  answer("Character.", correct = TRUE),
  answer("Boolean."),
  allow_retry = TRUE
)
```

```{r logical_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
gradethis::question('What data type is `FALSE`?',
  answer("Numeric."),
  answer("Character."),
  answer("Boolean.", correct = TRUE),
  allow_retry = TRUE
)
```

```{r numeric_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
gradethis::question('What data type is `42`?',
  answer("Numeric.", correct = TRUE),
  answer("Character."),
  answer("Boolean."),
  allow_retry = TRUE
)
```

<!-- End tutorial version -->
