### Combining data with operators

\index{operators}

Okay, now we know about data.

But data by itself is not especially useful. It just sits there until you *do something* to it. There are many ways of doing things to data, but some of the simplest are **operators**.

Operators *operate* on data. You may not have heard the name *operator* before, but you are already familiar with many common mathematical operators, such as `+` and `-` for adding and subtracting numbers.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise .callout-note appearance='simple' icon=false}
**Exercise (Try it yourself):** Try entering a number after the `>` in the Console (e.g. 1), then Enter, and see what happens.
:::

When you hit enter, the R interpreter reads in the line, evaluates it, and returns the answer. In this case, you entered 1, so the computer thinks 'Hey, it’s a 1! Wow, a one! The result of 1 is… drum roll, please… 1!' and returns the result of this **expression**, which is a one.

Cool! But not, I confess, particularly useful. Let’s fix that: next we'll add two numbers together. 

:::{.exercise .callout-note appearance='simple' icon=false}
At the prompt, enter two numbers separated by a plus sign, +

For example:

`> 1 + 1`

What do you get?
:::

(Note that I've left the Console's `>` prompt in the example code above, but I will leave it out in future examples.)
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Let's try using the addition operator, `+`, on some numbers to see what happens. Run this code chunk:
</p>
")
```

```{r operators_simple_addition, eval=is_tutorial, echo=F, exercise=TRUE}
1 + 1
```

```{r operators_simple_addition-check, eval=is_tutorial, echo=F}
grade_result(
  pass_if(~ identical(.result, 2), "Unsuprisingly, the + operator adds two numbers together.")
)
```

<!-- End tutorial version -->

Great! Let's move on and investigate operators in more depth...

### Operating on numbers

Heart surgeons operate on hearts, brain surgeons operate on brains. 
You will be operating on numbers... does that make you a data surgeon?

Here are some of the operators available to us in R:

| Operator  | Example  | Result  |
|:---------:|:--------:|:-------:|
| `+`       | `5 + 2`  | `7`     |
| `-`       | `5 - 2`  | `3`     |
| `*`       | `5 * 2`  | `10`    |
| `/`       | `5 / 2`  | `2.5`   |
| `^`       | `5 ^ 2`  | `25`    |
| `%%`      | `5 %% 2` | `1`     |

Some of these might seem obvious, while others might be unfamiliar. 
In this section's exercises we will go through them all and figure out what they do.



<!-- Book version -->
```{r, eval=is_book, echo=F, results = 'asis'}
opts_p <- c(
   "Adds two numbers together.",
   answer = "Subtracts one number from another.",
   "Multiplies two numbers together.",
   "Divides one number by another."
)
cat("
:::{.callout-note appearance='simple' icon=false}
**Exercise (the `-` operator):** In the R Console, type `5 - 2` and hit enter to run the line of code. (You probably have a good idea of what `-` does, but try changing the numbers just to make sure!)

What does `-` do?",
longmcq(opts_p),
"
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `-` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_subtraction, exercise=TRUE, eval=is_tutorial, echo=FALSE}
5 - 2
```

```{r operators_subtraction_quiz, echo=FALSE, eval=is_tutorial}
question("What does the - operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another.", correct = TRUE),
  allow_retry = TRUE
)
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
opts_p <- c(
  "Adds two numbers together.",
  "Subtracts one number from another.",
  answer  = "Multiplies two numbers together.",
  "Divides one number by another."
)
cat("
:::{.exercise .callout-note appearance='simple' icon=false}
**Exercise (the `*` operator):** In the R Console, type `3 * 2` and hit enter to run the line of code.

What does `*` do?",
longmcq(opts_p),
"
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `*` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_multiplication, eval=is_tutorial, exercise=TRUE, echo=FALSE}
3 * 2
```

```{r operators_multiplication_quiz, eval=is_tutorial, echo=FALSE}
question("What does the * operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another."),
  answer("Multiplies two numbers together.", correct = TRUE),
  answer("Divides one number by another."),
  allow_retry = TRUE
)
```


<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
opts_p <- c(
  "Adds two numbers together.",
  "Subtracts one number from another.",
  "Multiplies two numbers together.",
  answer  = "Divides one number by another."
)
cat("
:::{.exercise .callout-note appearance='simple' icon=false}
**Exercise (The `/` operator):** In the R Console, type `3 / 2` and hit enter to run the line of code.

What does `/` do? (Just to be sure, try some other numbers.)",
longmcq(opts_p),
"
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `/` operator does (feel free to play around with the numbers):
</p>
")
```

```{r operators_division, eval=is_tutorial, exercise=TRUE, echo=FALSE}
3 / 2
```

```{r operators_division_quiz, eval=is_tutorial, echo=FALSE}
question("What does the / operator do?",
  answer("Adds two numbers together."),
  answer("Subtracts one number from another."),
  answer("Multiplies two numbers together."),
  answer("Divides one number by another.", correct = TRUE),
  allow_retry = TRUE
)
```


<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
opts_p <- c(
  "Returns the modulus (the remainder after dividing one number by another).",
  answer = "Calculates an exponential (raises one number to the power of the second).",
  "Evaluates an equality (whether one number is equal to another).",
  "Evaluates an inequality (whether one number is bigger or smaller than another)."
)
cat("
:::{.exercise .callout-note appearance='simple' icon=false}
**Exercise (The `^` operator):**In the R Console, type `3 ^ 2` and hit enter to run the line of code.

What does `^` do? (Hint: Try some other numbers, like `2 ^ 3` or `16 ^ 0.5`)",
longmcq(opts_p),
"
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Run this code chunk to figure out what the `^` operator does (play around with the numbers to see what happens):
</p>
")
```


```{r operators_exponent, eval=is_tutorial, exercise=TRUE, echo=FALSE}
4 ^ 2
```

```{r operators_exponent_quiz, eval=is_tutorial, echo=FALSE}
question("What does the ^ operator do?",
  answer("Returns the modulus (the remainder after dividing one number by another)."),
  answer("Calculates an exponential (raises one number to the power of the second).", correct = TRUE),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```


### The `%%` operator

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
opts_p <- c(
  answer = "Returns the modulus (the integer remainder after dividing one number by another).",
  "Calculates an exponential (raises one number to the power of the second).",
  "Evaluates an equality (whether one number is equal to another).",
  "Evaluates an inequality (whether one number is bigger or smaller than another)."
)
cat("
:::{.exercise .callout-note appearance='simple' icon=false}
**Exercise (The `%` operator):** Next up, a slightly trickier one. Type `3 %% 2` and hit enter to run the line of code.

What does `%%` do? You will probably have to try some other numbers to figure this one out.

If you have difficulty, try also dividing the same numbers. E.g. try both 
`9 %% 4` and `9 / 4`.",
longmcq(opts_p),
"
:::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, eval=is_tutorial, echo=F, results = 'asis', echo=FALSE}
cat("<p class='comment3'>
Run this code chunk to figure out what the `%%` operator does (play around with the numbers to see what happens - for instance, try `4 %% 3` as well as `6 %% 3` and `3 %% 3`):
</p>
")
```

```{r operators_modulo, eval=is_tutorial, exercise=TRUE, echo=FALSE}
5 %% 3
```

```{r operators_modulo_quiz, eval=is_tutorial, echo=FALSE}
question("What does the %% operator do?",
  answer("Returns the modulus (the integer remainder after dividing one number by another).", correct = TRUE),
  answer("Calculates an exponential (raises one number to the power of the second."),
  answer("Evaluates an equality (whether one number is equal to another)."),
  answer("Evaluates an inequality (whether one number is bigger or smaller than another)."),
  allow_retry = TRUE
)
```


### Which operator goes first?

\index{precendence}
\index{order|see {precedence}}

Just like in normal math, we can do sums in R with multiple operators:

```r
3 + 5 / 5 * 3 ^ 2
```

In such a case, which operation do we do first?

Again, just like in regular math, some operations are always done before others. For example, all multiplication and division will be done before any addition or subtraction.

> **F.Y.I.**
>
> The order in which operators are calculated is known as *operator precedence*, and you can find the precedence of any operator here: [https://stat.ethz.ch/R-manual/R-devel/library/base/html/Syntax.html](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Syntax.html)

We can change the order of operations with parentheses: `(` and `)`. For example

```r
2 + 2 * 5 = 12
```

whereas

```r
(2 + 2) * 5 = 20
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
precendence_answer <- fitb(c("(3 + 5) / (5 * 2) ^ 2", "(3 + 5) / ((5 * 2) ^ 2)"), ignore_ws = TRUE, width = "20")
cat("
::::{.exercise .callout-note appearance='simple' icon=false}
**Exercise:** Modify this R code

`3 + 5 / 5 * 2 ^ 2`

so that it performs the calculation $\\frac{3 + 5}{(5 \\times 2) ^ 2}$.

Answer:",
fitb(
  c(
    "(3 + 5) / (5 * 2) ^ 2", 
    "(3 + 5) / ((5 * 2) ^ 2)"
    ), 
  ignore_ws = TRUE, 
  width = "20"
  ),
"
:::{.callout-tip collapse=true title='Expand for hint'}
When correct, you should get the answer 0.08.
:::
::::
")
```

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat("<p class='comment3'>
Modify the expression below so that it performs this calculation $\frac{3 + 5}{(5 \times 2) ^ 2}$ .
</p>
")
```

```{r operators_parentheses, eval=is_tutorial, exercise=TRUE, echo=FALSE}
3 + 5 / 5 * 2 ^ 2
```

```{r operators_parentheses-check, eval=is_tutorial, echo=F}
grade_result(
  pass_if(~ identical(.result, 0.08), "Looks like you're a parenthesis master!")
)
```

