
<!-- TODO: decide how to edit this with respect to earlier Boolean content - currently quite repetitive -->
### Boolean data revisted

At the start of this chapter we mentioned that there is a type of data in R, called Boolean data, that can have one of two values: `TRUE` or `FALSE`.

Then in the section on operators we saw that some operators (those that do comparisons) would give a `TRUE`/`FALSE` result even when the input data was not Boolean. For example,

```{r, include=TRUE, echo=TRUE}
2 == 2
```

because 2 is equal to 2 (i.e. this is a *true* expression).

The value `2` is an example of *numeric* data. However, the result `TRUE` is an example of **Boolean** data.

A Boolean value can only have one of two values (i.e. TRUE and FALSE). We often get a Boolean value when we do comparisons (e.g. checking if two values are equal, or if one is greater than the other).

Just as with numeric data, we can store a Boolean value in a variable, e.g. `d <- FALSE`.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat("
:::{.exercise}
Assign the value `TRUE` to a variable called `d`.
:::
")
```

<!-- End book version -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Assign the value `TRUE` to a variable called `d`.
</p>
')
```

```{r d_is_true, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="d_is_true-hint">
<p class="comment">
**Hint:** Try using `<-` to assign a Boolean value to a variable called `d`
</p>
</div>
')
```

```{r d_is_true-solution, eval=is_tutorial}
 d <- TRUE
```

```{r d_is_true-check, eval=is_tutorial}
grade_code(":)")
```

<!-- TODO: Maybe an exercise seeing if they can remember which operators were Boolean operators -->


### Combining pieces of data

So far we have looked at pieces of data by themselves:

```{r, echo=TRUE}
a <- 1
b <- "Hello!"
c <- TRUE
print(a)
print(b)
print(c)
```

But what about if we want to combine multiple pieces of data together?

R includes several types of *container* that can hold multiple pieces of data. We can then refer to that container by a single variable. For example, instead of the three variables above, we can create a **list** object that holds all three values:

```{r, echo=TRUE}
l <- list(1, "Hello!", TRUE)
l
```

We create a list using `list()`. Every value in the list goes inside the parentheses, separated by commas.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Create a list holding 4 values (in this order): 10, "z", FALSE, -0.1*10
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Create a list holding 4 values (in this order): 10, "z", FALSE, -0.1*10
</p>')
```

```{r create_a_list, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="create_a_list-hint">
<p class="comment">
**Hint:** No need to save this list to a variable...
</p>
</div>')
```

```{r create_a_list-solution, eval=is_tutorial}
list(10, "z", FALSE, -0.1*10)
```

```{r create_a_list-check, eval=is_tutorial}
grade_code(":)")
```

<!-- End book -->

### Vectors

Lists can hold data of different types. For example, our list `l` holds a number, a character string, and a Boolean value.

An alternative container in R is a **vector**. Vectors can also hold multiple pieces of data, but unlike a list, *all the data in a vector must be the same type of data*. For example, a vector could contain all numbers, or all characters, but not a mix of the two.

We create a vector with `c()`:

```{r, echo=TRUE}
c(1,2,3)
```

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Create a vector holding 3 character strings (in this order): "This", "is a", "vector!"
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Create a vector holding 3 character strings (in this order): "This", "is a", "vector!"
</p>')
```

```{r create_a_vector, eval=is_tutorial, exercise=TRUE}
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="create_a_vector">
<p class="comment">
**Hint:** No need to save this list to a variable...
</p>
</div>')
```

```{r create_a_vector-solution, eval=is_tutorial}
c("This", "is a", "vector!")
```

```{r create_a_vector-check, eval=is_tutorial}
grade_code(":)")
```

<!-- End tutorial -->

> You might be wondering what the numbers in square brackets at the start of each line in the output mean? E.g. `[1]`
>
> These tell us where abouts in the vector we are. The number indicates the position in the vector of the first element displayed on that line.
>
> For example, the `[1]` at the start of the line (before "Introduction") shows that "Introduction" is the first element in this vector.
### Operations on vectors

We can use operators on more complicated data structures just as we did on the simpler data types. For example, we can add 2 vectors together:

```{r, echo=TRUE}
v1 <- c(1,2,3)
v2 <- c(4,5,6)
v1 + v2
```
As you can see, the individual elements are added together.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
What happens if you add two vectors of different lengths? For example, run this code and see what happens:
:::
')
```

```{r, eval=is_book, results = FALSE}
v3 <- c(10, 20, 30, 40, 50)
v4 <- c(1, 2)
v3 + v4
```

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
What happens when you add two `v3` and `v4`?

* The 2 elements of v3 are added to the first 2 elements of v4, and the rest of v4 is unchanged.
* v4 is repeated to match the length of v3 before they are added.
* R guesses values for the remaining values of v3 so that the two vectors are the same length.
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
What happens if you add two vectors of different lengths?
</p>')
```

```{r vector_broadcasting, eval=is_tutorial, exercise=TRUE}
v3 <- c(10, 20, 30, 40, 50)
v4 <- c(1, 2)
v3 + v4
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="vector_broadcasting-hint">
<p class="comment">
**Hint:** Try changing the length of the shorter vector.
</p>
</div>')
```

```{r vector_broadcasting_quiz, eval=is_tutorial, echo=FALSE}
question("What happens when you add two vectors of different lengths?",
  answer("The 2 elements of v3 are added to the first 2 elements of v4, and the rest of v4 is unchanged."),
  answer("v4 is repeated to match the length of v3 before they are added.", correct=TRUE, message="This is called 'broadcasting', and is not what you might expect. In can cause bugs in R programs if the programmer (i.e. you) thinks that two vectors are the same length when they are not."),
  answer("R guesses values for the remaining values of v3 so that the two vectors are the same length."),
  allow_retry = TRUE
)
```

<!-- End tutorial -->
