The central component of everything we will be doing in R this semester is **data**. Even non-data science programs revolve around data. 

At a very basic level, a computer is just a fancy calculator that adds and subtracts numbers. Even things like words and pictures are stored inside a computer as numbers.

However, we often want to work work with data that is not numbers. For example, in the last section we were able to get R to print out the sentence *Hello, World!* To your computer that was just numbers flowing down wires as electrical signals. But the R programming language took care of converting our instruction into something your computer could understand.

This is the magic of programming languages! They allow us to write commands in (relatively) human-readable instructions, and then take care of translating that into the very unreadable numbers that computers work with.

R allows us to work with several "higher-level" types of data. These **data types** include:

* the **numeric** data type holds numbers such as `42` or `-12.5` or `0`. Unlike text, numbers are written *without* quotation marks around them. R sometimes refers to numbers as "integers" (if they are whole numbers) or "doubles" (if it is storing a number in a way that could handle decimals, since this (historically) takes up twice the amount of computer memory as an integer).

* the **character** data type holds text (i.e. letters, symbols, and the characters that represent numbers). We need to put the text inside quotation marks so that R knows where the text starts and ends: `"this is character data"`.

  * Note: in other programming languages this datatype is sometimes known as a "character string" or just a "string".

* the **Boolean** data type holds a value that is either `TRUE` or `FALSE`. (This is sometimes also referred to as the "logical" data type.)

<!-- Book version -->

```{r character-data-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  "numeric",
  answer = "character",
  "Boolean"
)
q_character_data <- mcq(answer_opts)
cat('::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** What data type is `"Introduction to Computing and Data for Scientists"`?',
q_character_data,
'
:::')
```

```{r numeric-data-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  answer = "numeric",
  "character",
  "Boolean"
)
q_numeric_data <- mcq(answer_opts)
cat('::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** What data type is `2`?',
q_numeric_data,
'
:::')
```

```{r boolean-data-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  "numeric",
  "character",
  answer = "Boolean"
)
q_boolean_data <- mcq(answer_opts)
answer_abbr_opts <- c("character", "double", answer = "logical")
q_r_bool_abbrv <- mcq(answer_abbr_opts)
cat(':::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** What data type is `FALSE`?', q_boolean_data,
'

*Follow-up:* In the RStudio Console, type in `typeof(FALSE)`, and hit the <ENTER> key to run this line of code. What do you get back?', 
'

::: {.info-box .callout-tip}
`typeof()` is another function. Whereas the `print()` function displayed simply displayed its own input, `typeof()` returns the data type of its input.
:::
::::')
```

```{r character-number-data-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  "numeric",
  answer = "character",
  "Boolean"
)
q_character_number_data <- mcq(answer_opts)
cat(':::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** What data type is `""5"`?', q_character_number_data,
'

::: {.info-box .callout-tip collapse="true" title="Explanation"}
As far as R is concerned, `5` and `"5"` are different types of data! `5` is a number whereas `"5"` is the character of the number.
:::
::::')
```

<!-- End book version -->

<!-- Tutorial version -->

```{r character_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
learnr::question('What data type is `"CDS 101"`?',
  answer("Numeric."),
  answer("Character.", correct = TRUE),
  answer("Boolean."),
  allow_retry = TRUE
)
```

```{r logical_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
learnr::question('What data type is `FALSE`?',
  answer("Numeric."),
  answer("Character."),
  answer("Boolean.", correct = TRUE),
  allow_retry = TRUE
)
```

```{r numeric_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
learnr::question('What data type is `42`?',
  answer("Numeric.", correct = TRUE),
  answer("Character."),
  answer("Boolean."),
  allow_retry = TRUE
)
```

```{r character_number_data_tutorial_quiz, eval=is_tutorial, echo=FALSE}
learnr::question('What data type is `"42"`?',
  answer("Numeric."),
  answer("Character.", correct = TRUE, message = 'Because "42" is inside quotation marks, it is not a numerical data but actually a string of two characters (the character "4" and the character "2"!'),
  answer("Boolean."),
  allow_retry = TRUE
)
```

<!-- End tutorial version -->
