### Tibbles

You can create a tibble using the `tibble()` function. The arguments of the function should be vectors that will form the columns of the table, for example:

```{r, echo=TRUE}
b <- c(TRUE, FALSE, TRUE)
z <- c("Anna", "Bob", "Carlos")

tibble(
  a = 1:3,
  4:6,
  student_name = z,
  b
)
```

Note that:

* if the vector is stored in a variable, the variable name will be the column name (e.g. column `b`)

* you can override the column name by using `new_name = the_vector` as we did for the 1st and 3rd columns

* if you do not, you may get an unusual column name, e.g. the 2nd column

<!-- Book version -->

<!-- TODO 
-->

<!-- End book -->

<!-- Tutorial version -->


```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Fill in the blanks to create a tibble with the column names: `student`, `grade`, `pass` (in that order)
</p>
')
```

```{r fjvsmmyysjpaxqmf, exercise = TRUE, echo=FALSE, eval=is_tutorial}
student <- c("Daphne", "Ed", "Frankie")
col_3 <- c(TRUE, TRUE, FALSE)
g <- c("A", "C", "F")

tibble(
  ____,
  ____,
  ____
  )
```

```{r fjvsmmyysjpaxqmf-hint-1, echo=FALSE, echo=FALSE, eval=is_tutorial}
# hint text
""
```

```{r fjvsmmyysjpaxqmf-hint-2, echo=FALSE, echo=FALSE, eval=is_tutorial}
# hint text
""
```

```{r fjvsmmyysjpaxqmf-solution, echo=FALSE, eval=is_tutorial}
# solution code
student <- c("Daphne", "Ed", "Frankie")
col_3 <- c(TRUE, TRUE, FALSE)
g <- c("A", "C", "F")

tibble(
  student,
  grade = g,
  pass = col_3
  )
```

```{r fjvsmmyysjpaxqmf-check, echo=FALSE, eval=is_tutorial}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->
