Before we begin to make our own graphs, we need to learn some terminology for describing the underlying data. 


<!-- 
TODO: New video.

Topics:
Tables // data frames (were)
Rows // observations (-> and what does each one represent)
Columns // variables (vs R variables)
Types of variable: continuous vs categorical, plus discrete, ordinal, count etc. Limitations of these distinctions. How to tell the type of data in R/RStudio.
-->


<iframe width="560" height="315" src="https://www.youtube.com/embed/zM5y3dZ-o6o?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://gmuedu-my.sharepoint.com/:b:/g/personal/dwhite34_gmu_edu/ETqhI64qDaFGja02KJaA_usBe0iEVyOhjM5kH7ENZswmsg?e=mg4B0G)


<!-- TODO: Written version. -->

We usually want to store numerical data in tables, just like you might do with spreadsheet software like Microsoft Excel or Google Sheets. For example, instead of the separate `x_coords` and `y_coords` vectors from the graph in the previous sections, we would often be working with these as separate columns of a single table:

| x_coords   | y_coords   |
|:----------:|:----------:|
| 1          | 2          |
| 3          | 4          |

: Table of our 2 points {#tbl-two-points}

Because the R programming language is designed for analyzing data, it comes with a built-in data structure for storing tables of data: the **dataframe**.

We will learn how to create dataframes later in this book. For now, I will just give you pre-existing dataframes that we can analyze. If you want to follow along with these examples in RStudio, you can run this line of code in the Console:

```{r, include=FALSE}
df <- data.frame(x_coords = c(1,3), y_coords = c(2,4))
```

That creates an R variable called `df` that holds the table of our two points. If we enter the name of the variable at the Console, R will print out the dataframe stored in the variable:

```{r}
df
```

As you can see, this contains the same data as the table above, albeit not formatted quite as nicely (the default formatting of dataframes in R is not particularly pretty, but that is another thing we will learn to do in the future!).

Our dataframe is broken down into various parts:

* Each *row* represents **an observation** of some thing. For example, the first row represents measurements taken from one thing, and the second row represents measurements taken from a different thing.

* Each *column* represents a different **variable**, or quantity that we are measuring.

  * Note that we are reusing the word *variable* here. We have already encountered one type of variable in the R programming language, as a way to store and reference objects and data in our computer.
  
    Now we have encountered a new use: variables as measurements from the real world, which exist as a columns in a table.
    
    These two concepts of a variable are different, even though we use the same word, "variable" to refer to each of them. I will be using "variable" for both of them throughout this book, and you will need to deduce from the context whether I am talking about a programming variable in R or a column variable that measures something in the real world. (If it could be ambiguous, I might refer to an "R variable" or a "column" to prevent confusion.)

  * The columns of an R dataframe are actually vectors! I.e. a dataframe is essentially just a bunch of vectors of the same length. This means that the rules that apply to a vector also apply to a dataframe column. For example, a dataframe column can only contain one type of data (e.g. just numbers, or just character strings), just like a vector.
  
    We can get a single column from a dataframe using the `$` operator after the dataframe's variable. For example, to refer to the `x_coords` column from our `df` dataframe, we could write:
    
    ```{r}
    df$x_coords
    ```

    And we get back a vector of the two values from that column!

We have already talked about the specific types of data that the R programming language is aware of, such as numbers, character strings, and Boolean values.

But there are also more abstract concepts of data that a column variable can contain. For example:

* *Continuous variables*: we say that a column variable is continuous if it can feasibly be any number. For example, measurements of the height of different humans is continuous, because you could be 200cm tall, or 201cm tall, or 200.5cm tall, or 200.75cm tall (etc.). The values between any two values are also valid values for a continuous variable to take on.

* *Discrete variables*: we say that a variable of numbers is discrete if it can only contain particular values. For example, the year is discrete, because it could be 2023 or 2024, but it cannot be 2023.5.

  * *Count variables* are a particular sub-type of discrete variable. As you might guess from the name, this is when the variable indicates how many there are of something.

* *Categorical variables* contain labels, text, or something other than numbers. For example, colors (red, blue, green, etc.) would be categorical values.

  * Categorical variables can be unordered, such as the color example, or...
  
  * They can be ordered, in which case we refer to them as *ordinal* variables. For example, if a variable of ratings can contain the values "good", "average", or "bad", then this would be considered an ordinal variable since there is an ordering to these values.

Here is how these different abstract types of variables could be concretely recorded in R in the columns of a dataframe (note that we have not yet encountered all of these R data types):

| Type of variable | Type of R data |
| :--------------: | :------------: |
| Continuous       | Numerical      |
| Discrete/count   | Numerical, integer (p. ??) |
| Categorical      | Numerical, character, factor (p. ??) |


<!-- TODO: some practice exercises:
* what is the observation in different tables.
* what is the type of data in different columns.
-->
