Another type of graph that you may encounter for two continuous variables is the line graph.

However, it is important to note that a line graph is not appropriate for many datasets. 

Here's why. When we connect points with a line, we are implying that there is some connection between those points. Typically we connect points in the direction of the x-axis, going from left to right. By drawing a line between, for example, the points (1,1) and (2,3) we are implying that a particular thing had the y-axis attribute equal to 1 when x=1, and then when $x=2$ that same thing had changed its y-axis attribute to a value of 3.

In other words, we should only draw a line graph to connect points if those points represent sequential observations of the same thing. A line graph shows us changes in values, whereas a scatter plot typically shows independently observed values of the same thing.

In would be bad, for example, to use a line graph to show the mass vs. height of Star Wars characters, because there is no reason to connect the point for Jabba the Hutt to the point for C-3PO[^1]. These rows in the dataset are completely separate things, not two observations of the same thing. If we compare such a line graph to the previous scatter plot that we made, we can clearly see that the line graph looks awful:

```{r, echo=FALSE}
par(mfrow = c(1, 2))

plot(x = starwars$height, y = starwars$mass)
plot(x = starwars$height, y = starwars$mass, type="l")

# Back to the original graphics device
par(mfrow = c(1, 1))
```

We typically use line plots when we are measuring something repeated over time. Time is then the sequential value on the x-axis (because time in graphs always goes from left to right). A dataset that contains repeated observations of the same thing over time is called a *time series*.

For example, here is a scatter plot and a line graph showing the body temperature of a beaver over time. All the body temperature measurements are from the same beaver, so this is a time series and it makes sense to connect the points. In fact, it is actually easier to understand the line graph than a scatter plot of the same data, because it is hard for our brain to mentally draw the connecting lines in the scatter plot.

```{r}
beaver_1_plus <- beaver1 %>%
  mutate(
    hour = time %/% 100,
    minutes = time %% 100,
    day_and_time = ddays(day-346) + dhours(hour) + dminutes(minutes)
  )
par(mfrow = c(1, 2))
plot(
  beaver_1_plus$day_and_time, 
  beaver_1_plus$temp, 
  xlab = "Time (seconds)", 
  ylab = "Body temperature (Celsius)"
  )
plot(
  beaver_1_plus$day_and_time, 
  beaver_1_plus$temp, 
  xlab = "Time (seconds)", 
  ylab = "Body temperature (Celsius)", 
  type = "l"
  )
par(mfrow = c(1, 1)) # Back to the original graphics device
```

We can create a line graph using the `plot()` function with the additional argument `type = "l"`, for example:

```r
plot(x = column_1, y = column_2, type = "l")
```

We supply the character `"l"` to the `type` parameter to indicate that we want **l**ines in this graph.


[^1]: Nobody suggest this to JJ Abrams...
