<!-- TODO: video 

topics:
- comparing the counts in difference categories

-->

We can also visualize the distribution of a categorical variable. We gain use a graph with bars, but whereas in a histogram we had to divide the x-axis into bins, we now have existing categories in the variable.

We can visualize the distribution by showing the number of values in each category (the *frequency* of each category).

To see how, let's pick the `sex` variable from the `starwars` dataframe. Here's what the first few rows looked like:

```{r, echo = FALSE}
knitr::kable(
starwars |>
  select(name, sex) |>
  head()
)
```

The first thing we need to do is count the number of values of each sex in the `sex` column. We can do this with the `table()` function:

```{r}
table(starwars$sex)
```

We can then use the `barplot()` function to make a bar graph of those frequencies:

```{r}
h <- table(starwars$sex)
barplot(h)
```

Note that saving the table to a variable (which we have called `h`) is optional - we could instead nest the `table()` function inside the `barplot()` function like this:

```r
barplot(table(starwars$sex))
```

:::{.callout-tip}
<!-- TODO: move this to data wrangling chapter -->
### Nesting vs sequential functions

When should you nest functions and when should you run them sequentially?

In general, nested functions can be harder to read. It's fine if you only have two or three, but nesting more than that results in a lot of parentheses and arguments that are hard to sort out.

We can instead run each function one at a time, and save the output in a new variable. However, if we only want to use that variable once (in the next function), then we end up creating a lot of variables that we don't really need.

In the next chapter we will learn about a new operator (called a *pipe*) that will allow us to run functions sequentially and avoid creating lots of intermediate variables.
:::
