How can we visualize a single column of data by itself? And why would we want to do this?

<!-- 
TODO: New video.

Topics:
- creation with hist()
- parameters: breaks for # of bins, and picking a good number
- interpretation: shape and center, outliers
- one or two scientific examples of why we care: pareto (wealth), height (bimodal)
-->

Let's use a dataset that is included in the `tidyverse` R package which contains information about different characters from the Star Wars franchise.

<!-- TODO: use a more scientific dataset than Star Wars -->

You can load the dataset in RStudio by loading the `tidyverse` package (which you installed in @sec-installing-packages) by running `library(tidyverse)`. The dataset will then be available to you as a dataframe stored in an R variable called `starwars`.

Here is an example of the data stored in the `starwars` dataframe:
```{r}
starwars
```

We might be interested in visualizing the heights of different Star Wars characters. Maybe we want to know if there are lots of tall characters and not many short ones.

A graph of a single variable shows us the **distribution** of that variable. A good type of graph to visualize the distibution of a single continuous variable (like the `height` column) is a **histogram**.

Here is how we can create a histogram of the `height` column with the `plot()` function:

```{r}
hist(x = starwars$height)
```

Note how:

* We used the `$` operator to select the `height` column as a vector of numbers to go on the x-axis.

* The x-axis of heights has been broken up into a series of bars which we call *bins*.

* The length of each bin is the number of observations (the "frequency") that fall into that bin (i.e. the number of Star Wars characters with a height in the range of that bin).

  * For example, the tallest bin catches the characters with a height between about 160-180 centimeters (cm). There seem to be about 28 characters in the dataset with a height in that range. (If you prefer the old Imperial measurement system, then 2.5cm is about 1 inch, so 100cm = 3'4", 150cm =  5', 200cm = 6'8", etc.)

R has guessed the number of bins for us. Sometimes you will want to customize this, and we can override the default by setting the `breaks` parameter of the `hist()` function. For example, we can force the histogram to contain just 5 bins:

```{r}
hist(x = starwars$height, breaks = 5)
```

R will recalculate the range of the new bins, and recount how many observations fall into each bin.

<!-- Modify for book and tutorial versions -->

**Exercise:** Find the tallest bin in the 5-bin histogram. What is the range of heights of Star Wars characters in this bin? Approximately how many characters are in this bin? Recall that one of the purposes of a graph is to see if the data fits our expectations - does this make sense as the most common height range to find in the Star Wars movies?

We can also create a histogram with a lot of bins, e.g. 100 bins:

```{r}
hist(x = starwars$height, breaks = 100)
```

Picking the right number of bins is a bit like Goldilocks tasting porridge: there's a sweet spot in the middle, but it's a bit subjective.

* If we have too few bins then we destroy any patterns in the variable.

* If we have two many bins then we see a jagged pattern caused by random variation rather than the true shape of the distribution.

Compare these three histograms with different numbers of bins:

```{r, echo=FALSE}
par(mfrow = c(1, 3))

hlarge <- hist(x = starwars$height, breaks = 2, main = "3 bins")
hmed <- hist(x = starwars$height, breaks = 9, main = "10 bins")
hsmall <- hist(x = starwars$height, breaks = 50, main = "50 bins")

# Back to the original graphics device
par(mfrow = c(1, 1))

```

* The middle histogram shows us that while humans (and similar-sized creatures) are predominant even a long time ago in galaxies far, far away, there is also a clump of shorter creatures with heights below 100cm.

* However, that shape is obscured in the plot on the left with just three bins. With too few bins, we destroy any fine grained shape in the distribution.

* It is equally hard to see the overall shape in the right-hand histogram, because there are so many bins that our eyes are distracted by the sharp transitions between neighboring bins. However, these patterns are caused by random variation, i.e. by accidentally having a few more Star Wars characters in one bin.

