### The pipe `%>%` operator

<iframe width="560" height="315" src="https://www.youtube.com/embed/SdrBENv3F-s?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1SCd5J-w1Z9E6vmT2LbmaeDZaE8pDA2p4)

As described in the video, the pipe operator takes the value on its left and inserts it as the first argument of the function on the right.

In other words:

```r
some_data %>% someFunction()
```

is equivalent to:

```r
someFunction(some_data)
```

If there are other arguments supplied to the function, they get "pushed back" so that the data piped in can claim the "first argument" spot:

```r
some_dataframe %>% someFunction(this_is_really_argument_2)
```

is the same as:

```r
someFunction(some_data, this_is_really_argument_2)
```

### Piping practice

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Exercise TBM
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Rewrite this code so that it uses a pipe to insert the `presidential` dataframe in the first argument position of `select`:

`select(presidential, name, party)`
</p>')
```

```{r kammbwbduyotylwz, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Your code here
____ %>%
  select(_____, _____)
```

(Note that we can put a new line after the `%>%` operator as above - R knows that there must be a right-hand side, so it treats both lines as the same line.)

```{r kammbwbduyotylwz-solution, eval=is_tutorial, echo=FALSE}
presidential %>%
  select(name, party)
```

```{r kammbwbduyotylwz-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->


**Why does this matter?**

The first argument of functions in the *tidyverse* is the dataframe. However, many functions also output a dataframe (as does a variable holding a dataframe, such as `presidential`). So we can just pipe from one function to another and build up a long chain of functions: i.e. a *pipe*:

```r
some_dataframe %>%
  select(some_columns) %>%
  some_other_function() %>%
  a_third_function(different_argument)
```

> Note that I have invented some made-up functions and variable names in the code above - this is called *pseudocode*.

