### The `select` function

The `select()` function can be used to pick certain columns of a dataset. The output of `select()` is a new dataframe containing just the columns that you specified.

**Ignore the video's instruction to follow along in RStudio**: you will try this function out on the next page of this tutorial instead.

<iframe width="560" height="315" src="https://www.youtube.com/embed/bfmMYUOWnls?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1DtuT-rtWs-i6MzWiYtX4HUvlqpIlZfUm)

### A simple `select()`

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Exercise TBM
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Select just the `name` column from the `presidential` dataset.
</p>')
```

```{r ques3, exercise=TRUE, eval=is_tutorial}
# Replace the blank with the name of the column
select(presidential, _____)
```

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<div id="ques3-hint">
<p class="comment">
**Hint:** Replace `column_name` with `name`
</p>
</div>')
```

```{r ques3-solution, eval=is_tutorial}
select(presidential, name)
```

```{r ques3-check, eval=is_tutorial}
grade_code("Nice work!")
```

<!-- End tutorial -->

### Selecting multiple columns

We can put as many column names as we want into the `select` function, separating each by a comma.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Exercise TBM
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Select both the `name` and `party` columns from the `presidential` dataset.
</p>')
```

```{r ullkyutezymdlivi, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Replace the blanks:
select(____, _____, _____)
```

```{r ullkyutezymdlivi-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"Fill in the 3 blanks with the name of the dataset and the two columns"
```

```{r ullkyutezymdlivi-hint-2, eval=is_tutorial, echo=FALSE}
# hint text
"Make sure that you have a comma between each argument, and that your function has opening and closing parentheses!"
```

```{r ullkyutezymdlivi-solution, eval=is_tutorial, echo=FALSE}
# solution code
select(presidential, name, party)
```

```{r ullkyutezymdlivi-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->


### Selecting a range

In R, the range operator, `:` (the colon punctuation symbol) indicates a range.

For example, this code create a vector of all the integer numbers in the range of 1-10:

```{r}
1:10
```

(We could also have written `c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)` as you learned in the first week, but the range operator is a convenient shorthand in this scenario.)

We can also use the range operator to indicate a range of columns inside `dplyr` functions such as `select`:

```r
select(presidential, name:end)
```

This selects all sequential columns from `name` to `end` (which in this case is `name`, `start`, and `end`).


<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Exercise TBM
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Using the range operator, select just the `start` and `end` columns.
</p>')
```

```{r uupnzgtdbacueimj, exercise = TRUE, eval=is_tutorial, echo=FALSE}
# Write your code here

```

```{r uupnzgtdbacueimj-solution, eval=is_tutorial, echo=FALSE}
select(presidential, start:end)
```

```{r uupnzgtdbacueimj-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

> Note:
>
> You combine ranges and individual column selections by separating them by commas, e.g.
>
> ```r
> select(presidential, name:start, party)
> ```

<!--
TODO
### Inverse selections
-->