### The `pivot_longer()` function

<iframe width="560" height="315" src="https://www.youtube.com/embed/B-OTLMCYYGQ?rel=0" frameborder="0" allowfullscreen></iframe>


<!--
TODO:
- Add new slides
-->
<!-- Slides: [PDF](https://drive.google.com/file/d/1CTxz65d4TfHhCAl3AaF1guNl3OvFABgm) -->


### Practice

Let's try the `pivot_longer()` function on the `presidential` dataset. We will reshape this dataset to convert the two data columns (`start` and `end`) into rows, with a names column that indicates the name of the original column, and a values column that holds the dates.

In other words, we want to convert the `presidential` dataframe:

| name | start | end | party |
| ---- | ----- | --- | ----- |
| Eisenhower | 1953-01-20 | 1961-01-20 | Republican |
| Kennedy    | 1961-01-20 | 1963-11-22 | Democratic |
| ... | ... | ... | ... |

into this:

| name | type_of_date | date | party |
| ---- | ------------ | ---- | ----- |
| Eisenhower | start | 1953-01-20 | Republican |
| Eisenhower | end | 1961-01-20 | Republican |
| Kennedy | start | 1961-01-20 | Democratic |
| Kennedy | end | 1963-11-22 | Democratic |
| ... | ... | ... | ... |

The data in both dataframes is the same, but we have changed the *shape* of the dataframe by converting columns into rows.

<!-- Book version -->

```{r, eval=is_book, echo=F, results = 'asis'}
cat('
:::{.exercise}
Exercise TBM
:::
')
```

<!-- End book -->

<!-- Tutorial version -->

```{r, eval=is_tutorial, echo=F, results = 'asis'}
cat('<p class="comment3">
Fill in the 4 blanks to achieve the reshaping operation described above.
</p>')
```

```{r xyryspqqovaxcgou, exercise = TRUE, eval=is_tutorial, echo=FALSE}
presidential %>%
  pivot_longer(cols = _____ | _____ , names_to="_____", values_to="______")
```

```{r xyryspqqovaxcgou-hint-1, eval=is_tutorial, echo=FALSE}
# hint text
"Replace the blanks with the names of the original columns and the two new columns."
```

```{r xyryspqqovaxcgou-hint-2, eval=is_tutorial, echo=FALSE}
# hint text
"List start before end in `pivot_longer()`: the order doesn't actually matter, but the grader doesn't know that :("
```

```{r xyryspqqovaxcgou-solution, eval=is_tutorial, echo=FALSE}
# solution code
presidential %>%
  pivot_longer(cols = start | end , names_to="type_of_date", values_to="date")
```

<!--
TODO:
- make sure the order of columns doesn't matter for the grader, and then remove hint #2
-->

```{r xyryspqqovaxcgou-check, eval=is_tutorial, echo=FALSE}
# check code
gradethis::grade_code()
```

<!-- End tutorial -->

