---
# execute:
#  freeze: auto
---

# Literate Programming

```{r setup}
#| results: "asis"
#| echo: false

source("prechapter.R")
insert_status_callout("incomplete")
```

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Add 'Quick reference' section at start of chapter with link to RMarkdown cheetsheet
")
```

Traditionally, if you needed to do science on a computer, you ran your analyses in one program, and then copied over any outputs you wanted into a separate document where you wrote your results. 

The separation of analysis and communication creates two main problems:

1. Copy errors: it is easy to forget exactly how or when results where generated when there are in a separate place on your computer. Further more, if you update the analyses half way through writing, you need to remember to copy over any of the new outputs into the written document and it's all to easy to miss something.

2. Discourages reproducibility: Typically just the final written document will be shared with other scientists, and not the underlying analyses that generated the results. Although the data/code repositories mentioned in the precious section can help to prevent this, it's an extra step.

One solution is an approach to programming with the fancy name of *literate programming* which just means combining writing and code in the same document. Then you *render* the document, which runs the code and generates the output of any analyses, and embed it within the text.

Having the analyses next to your written interpretation makes it easy for us to verify that we are writing about the analysis we think we are! It also means that your results will automatically be updated to the latest version whenever you compile the document. And finally, you can share this document and (theoretically) any other scientist should be able to compile it and hopefully understand the relationship between your analysis and results more easily.

Literate programming is not a new idea. It was first suggested by a prominent computer scientist called Donald Knuth in 1984[@knuth_literate_1984]. It didn't immediately take-off, because the early implementations were clunky and significantly harder to use than the usual method of writing code by itself in it's own file.

However, literate programming has taken strides forwards in the past decade. The R programming language has been a major part of this development, and includes one of the most developed systems for literate programming: **RMarkdown**, which appeared in 2014.

Prior to RMarkdown, we typically wrote R code in *source files* that could only contain code (you may encounter these files: they end with the .R extension). RMarkdown is an R package (i.e. an extension to the R language, see @sec-intro-r-functions) that allows sections of R code to be included in text documents.

Recently, an alternative to RMarkdown has appeared: **Quarto**.

Quarto is an attempt to expand RMarkdown to support other programming languages and types of input document. As we will see in the rest of this chapter, RMarkdown and Quarto are extremely similar, and you can achieve almost identical results with either system. 

This book is written using literate programming, using a combination of RMarkdown and Quarto files.

In the next few sections, we will go over how to create and edit RMarkdown and Quarto files. Most of the advice will be the same for each file type. Where the two systems differ, I will highlight this and show you both approaches. I will end the chapter with a more advanced discussion of the differences, in case you are planning a larger project and need to pick between the two (spoiler: it doesn't matter too much).

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Donald Knuth info box
")
```




```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: RMarkdown, Quarto, Formatting, Code chunks

Talk by  Yihui Xie: An anatomy of R Markdown with minimal applications: https://www.youtube.com/watch?v=fiy32LjgGUE
'it's possible to explain the basic ideas with minimal examples. In this talk, I'll show you the basic process of rendering R Markdown to HTML and LaTeX. After you learn the nuts and bolts, you may be able to create your own favorite output style, especially if you know CSS, JavaScript, and LaTeX techniques'
")
```

## RMarkdown and Quarto files

RMarkdown files are text files that end in the file extension `.Rmd`. For example, a file called `my_homework.Rmd` in an RMarkdown file. Quarto files are similar, but end with the extension `.qmd`.

We can create both types of files from within RStudio. Click on the *File* menu in the top left of RStudio^[or click the New File icon in the top toolbar], then pick the *New File* option, and then select either:

* *RMarkdown...* to create an RMarkdown file, or
* *Quarto document...* to create a Quarto file.

When you create new files in this way, RStudio will include some default contents to give an example of how each type of file works. (We will also see how to create empty files, with no default contents.)

Since the results are slightly different, let's look at each filetype in turn.

#### New RMarkdown files

When you create a new RMarkdown file, you will be presented with the following wizard:

{{< placeholder >}}
*Figure: RMarkdown creation wizard.*

You can accept the defaults in the wizard as shown above, and modify the rest later. Or you can enter a title, author name, and/or date now if you prefer. Click *OK* to create the file (which will contain some default contents).

{{< placeholder >}}
*Figure: Example default RMarkdown file.*

To create an empty file without any default contents, click the *Create Empty Document* button instead.

The file will automatically be opened in the Editor pane of RStudio.


#### New Quarto files

Quarto files come with a similar set-up wizard:

{{< placeholder >}}
*Figure: Quarto creation wizard.*

Again, it is fine to just click OK with the defaults, but you can put in a title and author name if you like.

I also like to deselect the *Editor* checkbox, but you can easily turn the visual editor on or off once you have created the file (@sec-visual-editor).

If you would like an empty Quarto document without any default contents, click the *Create Empty Document* button instead of *OK*.

The file will automatically be opened in the Editor pane of RStudio.

{{< placeholder >}}
*Figure: Side-by-side examples of default RMarkdown & Quarto files, with key features highlighted.*


## Rendering to an output

Note that the files you create are not saved when you create them. To save them, make sure the file is the current one open in the editor pane and either press the *Save* icon in the top toolbar, or *File > Save* in the top menu, or click anywhere in the open file and then press the {{< kbd Ctrl-S >}} keys at the same time.

To render the saved RMarkdown/Quarto files to an output:

* In RMarkdown, click the *Knit* button at the top of the open file in the Editor pane.
* In Quarto, click the *Render* button.

In each case, the default output filetype will be created, such as a PDF or a HTML file (a webpage). Later on we'll see change the type of output file you render to in @sec-yaml-header-section.

The rendering process will be shown in the bottom-right pane of RStudio while it happens. Sometimes a problem will occur which prevents the file from rendering. Common reasons for this are covered in the Frequently Encountered Problems section: @sec-common-render-issues.

Once rendering is finished, the file should open up automatically. A PDF will typically open in your computer's default PDF viewer. An HTML output file will open up in your browser. If the output file doesn't open automatically, you can open it manually by clicking on the file in the Files tab in the bottom-right pane of RStudio.

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Add try-it-yourself exercises: Try rendering the file.
")
```

## Text

If you look at Fig. X (rmd/qmd side-by-side), you can see that both Quarto and RMarkdown files are divided into two sections. At the top of each file is a *header* section, which begins and ends with a line of three dashes: `---`. This is where we can specify settings for the file, which we will learn more about in @sec-yaml-header-section.

Below that is the main section of the document. Here we can add text that will appear in our output when we render.

We can try adding text to a document to see what effect it will have on the output. On the left of Fig. X, we can see an RMarkdown and Quarto files with a brief header section (the same contents should work just as well in a Quarto file. Below that is a line of text that says `Hello, world!`. On the right are the outputs that we get when we render these files to HTML.

In the outputs, the text `Hello, world!` has been formatted as a paragraph of text.

{{< placeholder >}}
*Figure: Side-by-side examples of RMarkdown and output, and Quarto and output (hello world).*

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Add try-it-yourself exercises: (1) add 'Hello world'. (2) Remove header section and re-render.
")
```

What about if we want to format contents as something other than regular paragraph text? For example, you might want section headings (like this book contains), or *italics* or **bold text**. In a program like Microsoft Word or Google Docs, you can click a button to apply these formatting styles.

RMarkdown and Quarto files cannot contain this type of embedded formatting, because they are just text documents (i.e. the same thing as a file with a *.txt* extension). But we can specify that we want their output to contain formatting. We do this by including symbols in the input file that will be interpreted as formatting instructions when we render.

For example, we can create *italicized text* by putting the words inside a pair of asterisks `*italicized text*` or a pair of underscores `_italicized_text_`. The asterisks or underscores will not appear in the output, but the output will instead have the formatting you have specified.

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Add try-it-yourself exercises: (1) add italicization, (2) try mixing the symbols.
")
```

This method of applying formatting in a text document is called *Markdown*. You may have encountered something similar online, as websites like Reddit and Discord allow you to use symbols like this to apply formatting. It is also where *RMarkdown* files got their name from.

There are a lot of different symbols we can use to format text with Markdown.

As mentioned above, we can use a pair of asterisks or a pair of underscores to *italicize* text. In the same way, we can use double asterisks or underscores to get __bold__ text, and three to make text both ***italicized and bold***.

| Write this in your file  | To get this when you render |
|--------------------------|-----------------------------|
| `*italics* or _italics_` | *italics* or _italics_      |

````{=html}
<table>
  <thead>
    <tr>
      <th>Write this in your file</th>
      <th>To get this when you render</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td><span data-qmd="
```md
**bold** or __bold__

```"></span></td>
      <td>Regular output</td>
    </tr>
  </tbody>
</table>
````

:::: {.columns}

::: {.column width="50%"}

```md
*italics* or _italics_

**bold** or __bold__

_**bold and italicized**_
```

:::

::: {.column width="50%"}

*italics* or _italics_

**bold** or __bold__

_**bold and italicized**_

:::

::::


We create section headings by putting the hashtag/pound symbol `#` at the start of a line that contains the text of a heading. Two hashtags create a sub-heading (e.g. for a subsection), three hashtags create a sub-sub-heading, and so on. Note that there should be no spaces before the hashtags (it should be the first character on a line), but there should be a space between the hashtags and the first character of the text of the heading.

:::: {.columns}

::: {.column width="50%"}

```
# Heading
```

:::

::: {.column width="50%"}

# Heading {.unnumbered .unlisted}

:::

::::

:::: {.columns}

::: {.column width="50%"}

```
## Sub-heading
```

:::

::: {.column width="50%"}

## Sub-heading {.unnumbered .unlisted}

:::

::::

:::: {.columns}

::: {.column width="50%"}

```
### Sub-sub-heading
```

:::

::: {.column width="50%"}

### Sub-sub-heading {.unnumbered .unlisted}

:::

::::

Clickable links to webpages are created by putting the text of the link (what you will see) in square brackets, immediately followed by the URL (web address) of the webpage in parentheses. For example, here is a link to the Google homepage:

:::: {.columns}

::: {.column width="50%"}

```
[Link text](https://www.google.com)
```

:::

::: {.column width="50%"}

[Link text](https://www.google.com)

:::

::::

We can also create two types of list: unordered (i.e. bullet points) and ordered.

An unordered bullet point list is created by starting a line with an asterisk symbol `*` and then one or more spaces before the text of the item.

An ordered list is created by starting the line with a number followed by a period, and then one or more spaces before the text. You can also use roman numerals or letters of the alphabet to use different ordering symbols.

:::: {.columns}

::: {.column width="50%"}

```
* Bullet point item
* Second item
  * Nested list, indented by spaces
  * Another item
    - Another nested list, 
      with a different marker
```

becomes

* Bullet point item
* Second item
  * Nested list, indented by spaces
  * Another item
    - Another nested list, 
      using a hyphen marker

:::

::: {.column width="50%"}

```
1. An ordered list
2. Second item
   i. A nested list that uses 
       roman numerals
      a. A nested-nested list
   ii. Back to the roman numerals
   
       Second paragraph of item (ii).
```

becomes

1. An ordered list
2. Second item
   i. A nested list that uses 
       roman numerals
      a. A nested-nested list
   ii. Back to the roman numerals.
   
       Second paragraph of item (ii).

:::

::::

One important thing to note is that the number of spaces at the start of a list item matters.

Nested list items need to be preceded by the number of spaces to make them flush with the first character of the parent item's main text.

If a list item has multiple paragraphs, then (1) these paragraphs need to be separated by a blank line, and (2) all paragraphs after the first need to be indented with enough spaces so that their first character is flush with the first character of the first paragraph's main text.

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Add some kind of image that shows list indentation graphically.
")
```

::: {.callout-note}
### The importance of lines and spaces

In general, text on consecutive lines will be bundled up into one line when you render Markdown into another format. For example,

```md
This paragraph is written on two consecutive lines.
It will be compiled into a single paragraph because there is no blank line between them.

This will be a second paragraph, since there is a blank link between it and the previous text.
```

This can cause problems if you forget to put a blank line where you need it, e.g. before a heading line. For example,

```md
some regular text
### Important Heading
```

becomes

some regular text
### Important Heading {.unnumbered .unlisted}

Spaces can also affect formatting. 

If you forget a space between a heading's hashtags and its text, you will render it as normal text (with hashtags).

```md
###Heading Text
```

becomes

###Heading Text

The same thing can happen if you forget a space between a list symbol and the text of that list item.

Putting four spaces at the start of a line will cause it to be rendered like code:

```md
    Indented by four spaces.
```

becomes

    Indented by four spaces.

This has its uses if you need to render something like code, but is not appropriate for regular text. It is not only harder to read than properly formatted text, but it also will not be wrapped if the line is long (i.e. a long line will just run off the right side of the page instead of continuing onto the next line.)

:::

This brings us to the most important rule of writing and rendering Markdown based files:

::: {.callout-warning}

Always proofread the rendered output and check that the formatting is correct!

You can't always tell how the rendered output will look just from the contents of your RMarkdown/Quarto file. It's also very easy to make typos that will cause unintended formatting.

:::


## Code chunks

The great power of RMarkdown and Quarto documents is that we can also include sections of code within the document. These sections are called **code chunks**.

A code chunk begins and ends with a line of three backticks: \`\`\`. Note that a backtick is not a single quotation mark. On a US style keyboard it is the sumbol from the key in the top-left part of the keyboard.

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Better guide to where to find the backtick key
")
```

For example:

    ```
    code goes here
    ```

Backticks alone will format the code chunk like code, but you will not be able to run the code.

To indicate the programming language and thus to enable the code to be run, we need to include curly braces around the programming languages symbol at the end of the opening line of backticks. For example, this is an R code chunk:

````{verbatim}
```{r}
code goes here
```
````

A major reason for using RStudio to write Rmd/Qmd documents is that you can run code chunks from within the file editor tab. Either click on the green "Play" triangle in the top right, or click within the code chunk so that the cursor is flashing on a line of code and then simulatenously press the following keys: Ctrl + Shift + R (Windows/Linux) or Cmd + Option + R (Mac).

When you run a code chunk interactively from within the document, the code is sent to the Console to be executed, but any output is then displayed back below the code chunk. This allows us to preview what the output of a code chunk will look like without having to knit the entire document (which is a lot slower).

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
TODO: Figure (PDF) and/or gif (html) showing this
")
```

Because code in code chunks is sent to the Console, it runs in the same place (called an *Environment*) as non-code chunk code that you type directly into the Console itself. This also means that all code chunks share this same environment, and the results of one code chunk are available to another.

For example, if you create a R variable in one code chunk, a different code chunk can then use that variable.

There's just one problem: order matters. You have to run the code chunks in order. I.e. you would need to run the code chunk that creates the variable before you run the code chunk that references that variable. Otherwise the variable doesn't exist yet in the Environment and you will get an error.

There's a few ways that new R programmers run into errors due to this code chunk execution in the shared environment:

1. You mix up the order of the code chunks within your file. You can run them interactively out of order, and so avoid the error, but when you try to knit you get an error. (This is because the knitting process runs the code chunks from scratch (in a new Environment) in order that they exist in the file.)

2. You delete a code chunk that creates something another (not deleted) code chunk needs. That thing will stick around in your RStudio Environment, and so be availabe to any code chunks running interactively, but when you knit you get the same error as in the previous point (for the same reason).

3. You restart RStudio after a break and your most recently written code chunk doesn't work. This is because anything R creates in the Environment is temporary. It will only last for as long as R or RStudio is running on the computer. When you restart RStudio and reopen a Rmd/Qmd file, you will need to re-run any code chunks whose output you need. (It's often easiest to just rerun all the code chunks, which you can do by clicking the "Run" button in the top right of the file editor pane and then selecting the last option to "Run All".)

## Inline code

We can also include code within paragraphs of text. Let's say you calculated some important number in your analyses, and you want to discuss that number. You could type in the number by hand, but this is prone to errors where the calculated number changes for some reason (maybe you adjust the analysis's code, for example) but then you forget to update the number in the text.

We can instead just include the variable's value directly in the text using *inline code*, for example:

```{verbatim}
The answer to life, the universe, and everything is `r variable_name`.
```

Here, we use a single pair of backticks around the code to indicate where it starts an ends, and an `r` immediately after the first backtick to indicate that this is R code. When the document is rendered, the value assigned to the variable called `variable_name` will be rendered within the sentence instead of the code. For example (if `variable_name` = 42),

    My results showed that the answer to life, the universe, and everything is 42.

We can put more complicated expressions within the inline code backticks. For example, you could put a calculation to get the same result:

```{verbatim}
My results showed that the answer to life, the universe, and everything is `r 40 + 2`.
```
    
or we could get the date at the time that the document is rendered by using the `Sys.Date()` function that is part of the R programming language:

```{verbatim}
This document was created on `r Sys.Date()`.
```

which would be rendered as:

```
This document was created on `r Sys.Date()`.
```

<!--

Move to a more advanced chapter:


formatting:

* options
* names

* breaking up long lines of code (and style recommendations/guides)
* not running code (and optionally keeping syntax formatting - does this work in rmd like it does in GitHub flavored md)
* Code comments vs. self-documenting code vs. text explanations




## The header section {#sec-yaml-header-section}

* Outputting different file types

## Images & plots

Images (plus size and floating), and captions & references.

Plots produced by code chunks, and formatting.

## Tables

Markdown tables (& captions + references)

| Column 1     | Column 2 | Column 3 |
|--------------|:--------:|---------:|
| Left aligned | Centered | Right    |
| Cell 3       | Cell 4   | Cell 6   |

Tables produced by code chunks, and formatting.

## Math 


## Visual editor {#sec-visual-editor}

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
Refs:

* https://rstudio.github.io/visual-markdown-editing/
")
```

## Frequently encountered problems

```{r, echo=FALSE, results='asis'}
pdf_dev_comment("
Move these to the FEPs appendix.
")
```

### RStudio has forgotten my variables

### My formatting is wonky

### My document won't render {#sec-common-render-issues}

* Missing dependencies. R packages, Latex for PDF output.

* Deciphering error messages.
  * Run the code chunks interactively.
    * Make sure you didn't create a variable out of sequence, or delete a code chunk with a variable (try clearing the Environment)
  * Figuring out the offending code chunk.

* Non ASCII characters.

* Taking too long/crashes
  * Memory/ large outputs/slow code.
  

## RMarkdown versus Quarto

Aside from the differences covered above:

RMD has an extensive ecosystem. Quarto can be extended, but seems to be designed to contain most features within the main system (and because it's newer, doesn't have such an ecosystem).

Both can support multi-file projects, e.g. this book, but Quarto knits each file separately, whereas Rmd allows the files to be merged and subsequent files can access the output of code in previous files.

Personally I have found Quarto to be a little less flexible and more prescriptive, but not deal-breakingly so. It has some nice new features that are not standard in RMarkdown, and is being actively developed and improved (whereas RMd is relatively stable at this point). 

This book is written principally in Quarto.

Due to the similarities, its not too hard to switch between them, or even include RMarkdown files within Quarto files (which this book does).

## Topics covered elsewhere in this book

Links to RMarkdown/Quarto features covered elsewhere in this book:
* table formatting
* figure formatting
* paper templates (latex)
* outputs


## Useful resources

RMarkdown cheetsheet

Pandoc markdown guide: https://pandoc.org/MANUAL.html#pandocs-markdown

-->