\newpage
# Wrangling Data

In this chapter we will learn how to manipulate and transform tables of data.

## Wrangling overview

```{r, child=c('src/snippets/wrangling_overview.Rmd')}
```


## The `dplyr` package and the `tidverse`

```{r, child=c('src/snippets/wrangling_packages.Rmd')}
```


## The `presidential` Dataset

```{r, child=c('src/snippets/wrangling_dataset.Rmd')}
```


## Picking columns with `select()`

```{r, child=c('src/snippets/wrangling_select.Rmd')}
```


## Sorting with `arrange()`

```{r, child=c('src/snippets/wrangling_arrange.Rmd')}
```


## *Piping* data between functions

```{r, child=c('src/snippets/wrangling_pipes.Rmd')}
```


<!--
Note: Skipped over the slice function here
-->


## Boolean logic

```{r, child=c('src/snippets/wrangling_boolean_logic.Rmd')}
```



## Picking rows with `filter()`

```{r, child=c('src/snippets/wrangling_filter.Rmd')}
```


## Creating columns with `mutate()`

```{r, child=c('src/snippets/wrangling_mutate.Rmd')}
```


## Grouping and summarizing

```{r, child=c('src/snippets/wrangling_group_summarize.Rmd')}
```


## What is tidy data?

```{r, child=c('src/snippets/wrangling_tidy_data.Rmd')}
```


## Converting columns to rows

```{r, child=c('src/snippets/wrangling_pivot_longer.Rmd')}
```


## Turning rows to columns

```{r, child=c('src/snippets/wrangling_pivot_wider.Rmd')}
```


## Splitting and combining

```{r, child=c('src/snippets/wrangling_separate_unite.Rmd')}
```


## Other data wrangling functions

```{r, child=c('src/snippets/wrangling_other_functions.Rmd')}
```
