---
execute:
  freeze: true
---

# More Statistics

```{r setup}
#| echo: false
#| results: "asis"
source("prechapter.R")
suppressPackageStartupMessages(library(tidyverse))
insert_status_callout("draft")
```

## Statistical Inference Review

### Video overview

This video reviews the inference material covered in the first module.

<iframe width="560" height="315" src="https://www.youtube.com/embed/pVAxez7pA1M?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1uQMSTvCp1sWsH2OY1nLfUjpXaHKgRK6s/view?usp=sharing)


### Test your understanding

<!--
Other questions to add:
* What is observed statistic
* What is the null distribution
* interpreting the p-values meaning -> the probability of coming from the null
* alpha
-->

<!-- Book version -->

```{r response_vs_explanatory-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  "response = `treatment`, explanatory = `outcome`",
  answer = "response = `outcome`, explanatory = `treatment`",
  'response = `"survived"`, explanatory = `"drug"`',
  'response = `"survived"`, explanatory = `"placebo"`'
)
q_character_data <- mcq(answer_opts)
cat('::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** You give sick patients either a drug or a placebo, and you measure whether they die or survive. This is stored in 2 variables, `treatment` and `outcome`, which are both categorical. What are your explanatory and response variables?',
q_character_data,
'
:::')
```

```{r response_vs_explanatory-book-exercise, eval=is_book, echo=F, results='asis'}
answer_opts <- c(
  "0",
  "0.01",
  answer = "0.02",
  "0.05",
  "0.1",
  "0.2",
  "0.5",
  "1",
  "2",
  "5",
  "50"
)
q_character_data <- mcq(answer_opts)
cat('::: {.exercise .callout-note appearance="simple" icon=false}
**Exercise:** In a particular hypothesis test, we generate 10,000 permutations of our original data. 200 of them have a test statistic more extreme than our observed statistic. What is the p-value?',
q_character_data,
'
:::')
```

`r hide("Show explanation")`

The p-value is the probability that a value as or more extreme than the observed statistic came from the null distribution, which equivalent to the fraction of simulations more extreme than the observed statistic. This is 200/10,000 = 0.02

`r unhide()`


### Supplementary reading

These two chapters provide a good review of the original statistical inference material that we covered in the first inference module. You should review them before starting this week's assignment:

* Statistical Inference via Data Science by Chester Ismay and Albert Kim (available online at: https://moderndive.com/)
  * Chapter 7: Sampling - https://moderndive.com/7-sampling.html
  * Chapter 9: Hypothesis testing - https://moderndive.com/9-hypothesis-testing.html
  
On a lighter note, you can get an animated explanation of a hypothesis test with alpacas here: https://www.jwilber.me/permutationtest/

### Code review

In the first inference module, we used the `infer` package to simulate the null distribution.

We ran 10,000 random *permutations* to shuffle up the values of the response and explanatory variables. This random shuffling ensures that we simulate a distribution where there is no relationship between the two variables.

```r
<DATAFRAME> %>%
  specify(<RESPONSE> ~ <EXPLANATORY>) %>%
  hypothesize(null = "independence") %>%
  generate(10000, type = "permute") %>%
  calculate(stat = "...", order = c(..., ...))
```

* The `specify` function defines the response and explanatory variables.
* The `hypothesize` function declares what the null hypothesis is between those variables. `"independence"` means that there is no relationship between them.
* The `generate` function is where the simulations are actually run: we specify how many, and how to shuffle the data (in this case, using permutations).
* The `calculate` functions takes the simulations and calculates a *test statistic*: i.e. a single value for each one (e.g. the difference in proportions, or the difference in means).

<!--
### Test how much you remember!
-->



## One-sided vs. two-sided tests

### Video overview

<iframe width="560" height="315" src="https://www.youtube.com/embed/krwRh7G6FPI?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1EqzByiMEvi9V9MfPQ9XsLT-wcSkU69Fq/view?usp=sharing)


<!-- ### 1 or 2? -->

<!-- <!-- -->
<!-- Questions to add: -->
<!-- * Writing alternative hypotheses for 1 and 2 sides -->
<!-- * Which parameter in code specifies order of diff calculation -> order -->
<!-- * Examples of whether to use 1 or 2 sided. -->
<!-- * Calculating p-value -->
<!--   * 1 sided left and right -->
<!--   * 2 sided -->
<!-- --> -->

<!-- ```{r alt_hypothesis_specification-quiz, eval=is_tutorial} -->
<!-- question( -->
<!--   "You want to conduct a *two-sided* hypothesis test on data from an experiment comparing whether a patient survive a disease after receiving a drug or a placebo. What is your alternative hypothesis?", -->
<!--   answer("The difference in proportions of patients who survive is significantly different between drug and placebo groups.", correct = TRUE, message = "In a two-sided test, the alternative hypothesis is that there is a difference, regardless of direction (i.e. whether the difference is positive or negative is unimportant)."), -->
<!--   answer("The difference in means of patients who survive is significantly different between drug and placebo groups."), -->
<!--   answer("The difference in means of patients who survive is significantly greater in the drug group than the placebo group."), -->
<!--   answer("The difference in proportions of patients who survive is significantly greater in the drug group than the placebo group."), -->
<!--   allow_retry = TRUE -->
<!-- ) -->
<!-- ``` -->


## Effect sizes

### Video overview

<iframe width="560" height="315" src="https://www.youtube.com/embed/o7oJPQMUS6A?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/1Ysg-oR6FLpuJ2OQ9ArzPNLvakImMgpBP/view?usp=sharing)

<!--
### Question time

-->

<!--
## Resampling with permutations versus bootstraps

Coming soon.
-->

## Confidence intervals

### Video overview

<iframe width="560" height="315" src="https://www.youtube.com/embed/iDxp6DHMHcc?rel=0" frameborder="0" allowfullscreen></iframe>

Slides: [PDF](https://drive.google.com/file/d/11eyvzCjS_M-owi_6Ry3WH0_5sqk5Y009/view?usp=sharing)


<!-- ### Questions -->

<!-- <!-- -->
<!-- Extra questions? -->
<!-- --> -->

<!-- ```{r permutations_vs_boostrap_1_quiz, eval=is_tutorial} -->
<!-- question( -->
<!--   "You want to conduct a hypothesis test using the `infer` package in R. How should you shuffle the data to generate a null distribution?", -->
<!--   answer("Bootstraps"), -->
<!--   answer("Permutations", correct = TRUE, message = "A permutation shuffles up the values of the explanatory and response variables, and so simulates a world in which there is no relationship between the two (i.e. the null hypothesis)."), -->
<!--   answer("It doesn't matter"), -->
<!--   allow_retry = TRUE -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r permutations_vs_boostrap_2_quiz, eval=is_tutorial} -->
<!-- question( -->
<!--   "You want to calculate the confidence interval around a statistic (e.g. the mean) using the `infer` package in R. How can you approximate resampling the original population?", -->
<!--   answer("Bootstraps", correct = TRUE, message = "By taking random samples of rows from the original sample (*bootstrap resampling*), we can approximate resampling the original population."), -->
<!--   answer("It doesn't matter"), -->
<!--   answer("Permutations"), -->
<!--   allow_retry = TRUE -->
<!-- ) -->
<!-- ``` -->

<!-- TODO
Bonferoni correction
-->

## Further reading

If you wish to do further reading on *bootstrapping* and *confidence intervals*, here are some online book chapters on the topic. (They cover the same material, so pick the one that makes the most sense to you.)

* Statistical Inference via Data Science by Chester Ismay and Albert Kim
  * Chapter 8: Bootstrapping and Confidence Intervals - https://moderndive.com/8-confidence-intervals.html
  
* Statistical Modeling by Daniel Kaplan
  * Chapter 5: Confidence Intervals - https://dtkaplan.github.io/SM2-bookdown/confidence-intervals.html

